Actomyosin remodeling regulates biomineral formation, growth and morphology during eukaryote skeletogenesis

Version 1

Abstract

Abstract

Biomineralization had apparently evolved independently in different phyla, using distinct minerals, organic scaffolds and gene regulatory networks (GRNs). However, diverse eukaryotes from unicellular organisms, through echinoderms to vertebrates, use the actomyosin network during biomineralization. Specifically, the actomyosin remodeling protein, Rho-associated coiled-coil kinase (ROCK) regulates cell differentiation and gene expression in vertebrates’ biomineralizing cells, yet, little is known on ROCK’s role in invertebrates’ biomineralization. Here we reveal that ROCK controls the formation, growth and morphology of the calcite spicules in the sea urchin larva. ROCK expression is elevated in the sea urchin skeletogenic cells downstream of the Vascular Endothelial Growth Factor (VEGF) signling. ROCK inhibition impairs the organization of F-actin around the spicules, disrupts skeletogenic gene expression and leads to skeletal loss. ROCK inhibition after spicule formation reduces spicule elongation rate and induces ectopic spicule branching. Reduced skeletal growth and enhanced branching are also observed under direct perturbations of the actomyosin network. Similar skeletogenic phenotypes are observed when ROCK is inhibited in a skeletogenic cell culture, indicating that these phenotypes are due to ROCK activity specifically in the skeletogenic cells. We propose that the actomyosin network was employed independently, downstream of distinct GRNs, to regulate biomineral growth and morphology across Eukaryotes.

Introduction

Biomineralization is the process in which organisms from the five kingdoms of life, use minerals to produce shells, skeletons and teeth that protect and support them (1–4). Recent studies suggest that biomineralization evolved by the phylum specific co-option of ancestral gene regulatory networks (GRNs) that drove the construction of distinct organic scaffolds (4–9) and by the evolution of specialized sets of biomineralization proteins (3, 7, 10-13). This explanation is in line with the dissimilar GRNs and biomineralization proteins that drive biomineralization in different phyla (5, 12, 14). Yet, there are common cellular processes required for mineralization in highly diverse phyla (4), suggesting that distinct upstream GRNs might have recruited common cellular and molecular mechanisms to drive biomineralization.

The actomyosin network was shown to play a role in biomineralization in various Eukaryote models, from unicellular organisms to vertebrates’ bones and teeth (15–22). There is a tight association between actin filaments and the biomineralization compartment in diatoms (19) and a highly dynamic actin organization that forms the biomineralization compartment in foraminifera (17). Perturbations of actin polymerization result with severely deformed shells and inhibition of shell shedding in coccolithophores (15, 16) and diatoms (18). In vertebrates, Rho GTPases and Rho56 associated coiled-coil kinase (ROCK), regulate chondrocytes, osteoblasts and odontoblasts differentiation and affect genes expression in these biomineralizing cells (20–22). However, the roles of actomyosin remodeling in controlling mineral deposition and shaping biomineral morphology, are still unclear.

Sea urchin larval skeletogenesis provides an attractive model to study the role of the actomyosin network in biomineralization. Sea urchin larval skeletons are made of two frameworks of interconnected calcite rods, termed “spicules” that are generated by the skeletogenic cells (6, 23). To make the spicules, the skeletogenic cells form a ring with two lateral skeletogenic clusters and fuse through their filopodia forming a pseudopodia cable that links them into a syncytium (24, 25). The mineral is concentrated in the form of amorphous calcium carbonate (ACC) inside intracellular vesicles (26, 27). The vesicles are secreted into the biomineralization compartment generated in the skeletogenic cell clusters forming two triradiate spicules 68 (Fig. 7A, (26, 28)). The tubular biomineralization compartment, also called the spicule cavity, elongates within the pseudopodia cable by localized mineral deposition at the tip of each rod (29).

The GRN that controls sea urchin skeletogenesis is known in great detail and is highly similar to the GRN that controls vertebrates’ vascularization, suggesting a common evolutionary origin of these two tubulogenesis programs (6, 23, 30). As the spicules elongate, the expression of key regulatory and biomineralization related genes becomes restricted to the skeletogenic cells proximal to the growing tips, possibly to promote mineral deposition at these sites (31–33). Localized gene expression is regulated by signaling cues such as the Vascular Endothelial Growth Factor (VEGF) signaling (31, 32). However, how the skeletogenic GRN drives spicule formation and localized mineral deposition at the tips of the rods is poorly understood.

Previous works suggest that the actomyosin network is essential for sea urchin skeletal growth. Calcium bearing vesicles perform an active diffusion motion in the skeletogenic cells with a diffusion length that inversely correlates with the strength and activity of the actomyosin network (34). Actin filaments are formed around the spicule (34, 35) and F84 actin signal is enriched at the tips of the growing skeletal rods in skeletogenic cell culture (36). Genetic and pharmacological perturbations of the GTPase, CDC42, prevent the formation of filopodia in the skeletogenic cells and inhibit spicule formation and elongation (37). Pharmacological perturbations of ROCK prevent spicule formation (38) and genetic perturbations of Rhogap24l/2 result with ectopic spicule splitting (6). Despite these phenotypic characterizations, little is known on the regulation and role of the actomyosin network and remodeling proteins in localized mineral deposition.

Here we study the role of ROCK and the actomyosin network in the sea urchin Paracentraorus lividus (P. lividus). Our findings reveal the critical role of ROCK and the actomyosin network in multiple aspects of sea urchin biomineralization, suggesting a common use of these factors in Eukaryote biomineralization, downstream of distinct GRNs.

Results

ROCK is enriched in the skeletogenic cells depending on VEGF signaling

We sought to study the spatial expression 98 of the ROCK protein and its regulation by VEGF signaling, a prominent regulator of sea urchin skeletogenesis (6, 30). The sequence of ROCK and especially, its functional domains are highly conserved between human and sea urchin (Fig. S1, (39, 40)). We therefore used a commercial antibody that recognizes human ROCK protein to test ROCK expression in P. lividus embryos (Fig. S2). We first used western blot to detect ROCK expression in control and under VEGFR inhibition at the time of spicule formation and during skeletal elongation (∼22 hours post fertilization [hpf], 27hpf and 33hpf, Fig. S2). The antibody detected a ∼150kDa protein (Fig. S2A) that is the predicted size for P. lividus ROCK protein (153kDa). VEGFR inhibition marginally increased the overall level of ROCK at 22hpf, but did not affect it at 27hpf and 33hpf (Fig. S2B). Yet, this measurement was done on proteins extracted from whole embryos, and we wanted to study the spatial expression of ROCK and specifically, its regulation in the skeletogenic cells.

We studied the spatial distribution of ROCK protein using ROCK antibody and the skeletal cell marker, 6a9, that marks the skeletogenic syncytium membrane ((41), Fig. 1). We quantified ROCK signal in the skeletogenic cells compared to the neighboring ectodermal cells in control embryos and under VEGFR inhibition, at the three time points mentioned above. At 22hpf, ROCK expression in the skeletogenic cells is mildly enriched compared to the neighboring ectodermal cells and this enrichment is not observed under VEGFR inhibition (Fig. 1A-F). At 27hpf, ROCK enrichment in the skeletogenic cells increases, but similar enrichment is observed under VEGFR inhibition (Fig. 1G-L). At 33hpf, ROCK enrichment in the skeletogenic cells is most apparent (Fig. 1M-O) and depends on VEGF signaling (Fig. 1P-S). At both 27hpf and 33hpf in control embryos, ROCK seems localized near the skeletogenic cell membranes, which further supports its activation in these cells, since ROCK activation leads to its localization to the cell membranes in other systems (42). Overall, this data demonstrate that ROCK expression is elevated in the skeletogenic cells and this enrichment strengthens with skeletal elongation and depends on VEGF signaling.

ROCK enrichment in the skeletogenic cells increases with time and depends on VEGF signaling

A-R, ROCK immunostaining at different times points in control and under VEGFR inhibition. In each condition and time point, the left image shows ROCK immunostaining and the skeletogenic cell marker, 6a9 (A, D, G, J, M and P). The middle image shows ROCK immunostaining alone in the whole embryo (B, E, H, K, N and Q). The right image shows the enlargement of the white rectangle region of the middle image (C, F, I, L, O, R). Scale bar in whole embryo is 50µm and in enlargements is 10µm. E – ectoderm, SC – skeletogenic cells. S, quantification of the ratio between ROCK signal/area in the skeletogenic cells compared to the ectodermal cells (see methods for details). Each box plot shows the average marked in x, the median, the first and the third quartiles (edges of boxes) and all experimental measurement (dots). Experiments were performed in three independent biological replicates and in each condition at least 33 embryos were measured. Statistical significance was measured using paired 2-tailed t-test where, * indicates P<0.05, ** indicates P<0.005, and *** indicates P<0.0005.

ROCK activity in the skeletogenic cells controls spicule initiation, growth and morphology

Next, we studied the role of ROCK in sea 128 urchin skeletogenesis using genetic perturbations. We downregulated the expression of ROCK by the injection of two different translation morpholino anti-sense oligonucleotides (MASOs). ROCK MASO-1 matches the start of translation region, ROCK MASO-2 matches the 5’ region upstream the start of translation, and random MASO was used as a control. Embryos injected with either ROCK MASO-1 or MASO-2 show strongly reduced skeletons (Fig. 2C, D) and embryos injected with ROCK MASO-2 show an additional phenotype of severely branched skeleton (Fig. 2B). The additional branching phenotype and the larger percentage of affected embryos in ROCK MASO-2 are probably due to its higher efficiency resulting from its more 5’ location relative to translation start codon (Fig. 2E). Thus, the genetic perturbations of ROCK expression indicate that sea urchin ROCK is important for skeletal elongation and normal branching.

To elucidate the dependence of ROCK phenotypes on ROCK activity level and identify its effect on different stages of skeletogenesis, we tested the skeletogenic phenotypes of different concentrations of ROCK inhibitor, Y27632, applied before or after spicule initiation (Fig. S3A). Y27632 binds to the ATP site of ROCK’s kinase domain and prevents its activity (43). The amino-acids to which Y27632 binds, are conserved in the sea urchin ROCK protein, supporting its specific inhibition of this protein (Fig. S1C). Continuous ROCK inhibition, beginning at egg fertilization, using 80μM Y27632, did not affect skeletogenic cell migration but eliminated skeletal formation at 27hpf (Fig. SB, C). At two days post fertilization (dpf) the majority of the embryos completely lack skeleton and the rest show only spicule initiations, yet their overall embryonic development seems normal (Fig. 2G, S3I, in agreement with (38)). Similar results are observed when adding the inhibitor in this concentration at 20hpf (Fig. 2I, S3I). Yet, the addition of 80μM of the inhibitor at 25hpf, after spicule initiation, results in a majority of embryos with severe ectopic spicule branching, which resembles the branching phenotype observed in ROCK MASO-2 (Fig. 2H, I). Reducing the concentration of the added inhibitor leads to milder branching and elongation phenotypes (Fig. S3F, H, K). However, even at the lowest inhibitor concentration, we detect a reduction of skeletal growth rate and ectopic branching a few hours after the addition of ROCK inhibitor (Fig. S4, 30μM). Washing the inhibitor after 25hpf results in partial recovery of skeletogenesis, with normal or mildly delayed skeletons (Fig. S3E, K). Overall, our genetic and pharmacological perturbations indicate that ROCK activity is necessary for spicule formation, skeletal elongation and normal branching pattern.

ROCK activity is essential for spicule formation, normal elongation and branching in whole embryos and in skeletogenic cultures

A-E, genetic perturbation of ROCK translation using two different MASOs observed at 2dpf. A, control embryo injected with Random MASO. B, embryo injected with ROCK MO-2 shows ectopic spicule branching. C, D, embryos injected with ROCK MO-1 or MO-2 show spicule initiations. E, summary of MASO injection phenotypes based on 4-6 independent biological replicates. F-I, pharmacological perturbations of ROCK activity using 80µM of the inhibitor Y27632 observed at 2dpf. F, representative control embryo with normal skeletal rods, B, body; AL, anterolateral; PO, post-oral and MV, midventral. G, complete skeletal loss in embryo treated continuously with ROCK inhibitor. J, reduced skeletal growth and enhanced ectopic branching in embryo where ROCK inhibitor was added at 25hpf. I, summary of perturbation phenotypes based on 3-8 biological replicates for each treatment. See additional phenotypes and summary in Fig. S3 and Table S1. J-M, representative spicules from skeletogenic cell cultures in control and under 30µM Y27632 at 72hpf. J, linear spicule in control culture, K, Y27632 addition at 16hpf, before spicule initiation, completely blocks spiculogenesis. L, Y27632 addition after spicule initiation, at 48hpf, reduces spicule elongation, M, and enhances branching, N, Quantification of spicule length in control and ROCK inhibition (>48hpf) at 72hpf. *P < 0.05, **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment, except from 120µm that was done in two biological repeats. Scale bars are 50µm. In J, L, M, the numbers at the bottom indicate the number of spicules that show this phenotypes (left) over all observed spicules (right).

To test if ROCK skeletogenic phenotypes are due to its activity specifically in the skeletogenic cells, we inhibited ROCK in a culture of isolated skeletogenic cells (Fig. 2J-N). The addition of ROCK inhibitor to the skeletogenic cell culture at 16hpf, before the spicules form, completely abolished spicule formation (Fig. 2K). The addition of the inhibitor at 48hpf, after spicule formation, resulted with significantly shorter spicules (Fig. 2L, N) and increased branching compared to control spicules (Fig. 2M). The reduction of the spicule length becomes more significant with increasing concentration of the ROCK inhibitor (Fig. 2N). Overall, ROCK skeletogenic phenotypes in the isolated skeletogenic cell culture are similar to its phenotypes in whole embryos, verifying that it’s specific activity in the skeletogenic cells is essential for spicule formation, normal elongation and prevention of branching.

ROCK inhibition reduces spicule volume, surface area and total length, but not thickness (SR-µCT)

To quantify the effect of ROCK inhibition on mineral deposition and skeletal shape we used synchrotron radiation micro-computed tomography (SR-µCT) that enables three-dimensional visualization and quantification of spicule morphology at 2 and 3dpf (Figs. 3, S5, see methods for details). The aberrant morphology of ROCK inhibited spicules is visible at both developmental time points (Fig. 3A). The ectopic branching occurs both at the tips where tip-splitting is detected (arrowheads, Fig. 3A), as well as in the back, where small spicule extensions, or mineral ‘dripping’ are observed (arrow at 48hpf, Fig. 3A). This implies that the regulation of mineral deposition is perturbed both at the tip and in the back of the spicules under ROCK inhibition.

SR-µCT studies of ROCK inhibited spicuels show reduction in skeletal volume, surface area and total length, but not thickness and 2-pulse calcein shows loss of tip-dominance

A, Exemplary 3D-renderings of control spicules (top, blue) and spicules where 40µM of ROCK inhibitor was added at 25hpf (bottom, orange), dissected at 48hpf and 72hpf. Arrowheads point to tip splitting and arrow at 48hpf point to spicule dripping at the back. B, Spicule volume vs. area for control and ROCK-inhibited spicules at 48 and 72 hpf. Each data point represents a single spicule. C-D, Frequency distributions for volume and surface area of control and ROCK-inhibited spicules at 48 hpf and 72 hpf (left and right violin plots, respectively). E, Spicules’ total branch length and average thickness for control and ROCK-inhibited spicules at 48 and 72 hpf. F-G, Frequency distributions for spicule lengths and thickness of control and ROCK-inhibited spicules dissected at 48hpf and 72hpf. H, I, Calcein two-pulses experiment. Embryos were exposed to blue calcein from 29-30hpf, to green calcein from 44-45hpf and stained with FM4-64 membrane marker (red) prior to image acquisition at 48hpf. H, Control embryo, G, Embryo where 30µM of Y27632 was added at 25hpf. The experiments were done in three biological replicates and the numbers at the bottom indicate the number of embryos that show this phenotype out of all embryos scored.

When compared to control spicules, ROCK inhibition leads to a ∼2.5-fold reduction of the spicule volume and surface area in both 2 and 3dpf (Fig. 3B-D, see Table S2 for measured values and standard deviations and S3 for statistical analyses). These SR-µCT measurements show that the rate of mineral deposition is significantly reduced under ROCK inhibition.

To understand the specific effect of ROCK inhibition on spicule elongation vs. thickening, we compared the total length and the mean thickness of the spicules between control and ROCK inhibition (Fig. 3E-G). The total length of control spicules was on average about 2.5-times longer than the length under ROCK inhibition in both 2dpf and 3dpf (Fig. 4E, F, tables S2 and S3). In contrast to its effect on the spicule length, ROCK inhibition caused a minor increase of the spicule thickness at 2dpf and at 3dpf the thickness was not affected by the inhibition (Fig. 3E, G, Tables S2, S3). These SR-µCT measurements indicate that ROCK activity plays a crucial role in controlling the rate of mineral deposition at the tips of the rods and in spicule elongation, and is not affecting spicule thickening.

Actin polymerization and myosin activation affect skeletal growth and branching

A-F, representative embryos showing the effect of actomyosin perturbations at 2df. A, control embryo B, embryo treated with skeletogenic cell culture were, C, embryo treated with 2μM Blebbistatin >20hpf, D, embryo treated with 2nM Latrunculin-A and 1.5μM Blebbistatin after 25hpf, E, embryo treated with 2nM Latrunculin-A after 20hpf, F, embryo treated with 2nM Latrunculin-A after 25hpf, arrow pointing to the additional spicule rod. G, Statistics of Latrunculin-A and Blebbistatin phenotypes, color code of phenotype is indicated in the representative images. Error bars indicate standard deviation. All treatments were conducted in at least three biological replicates and exact number of replicates and scored embryos are provided in table S4. H-J, representative spicules recorded at 72hpf from H, control skeletogenic cell culture, I, skeletogenic cell culture were 2nM Latrunculin-A was added at 48hpf and J, skeletogenic cell culture were 2μM Blebbistatin was added at 48hpf. K, Quantification of spicule length in the different treatments at 72hpf **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment. Scale bars are 50μm.

ROCK activity does not control mineral uptake but is required for “tip-dominance”

To test the effect of ROCK inhibition on mineral intake, we used the fluorescent chromophore, calcein that binds to divalent ions including Ca+2 and is widely used to track the calcium pathway in biomineralization (6, 28, 33, 34). We measured the number of calcein stained vesicles per area as well as the number of pixels of calcein signal per area in the skeletogenic cells, in control and under ROCK inhibition at 1dpf and 2dpf (Fig. S6A-E). Continuous ROCK inhibition does not change the number of calcium bearing vesicles at 1dpf, but the number of calcein stained pixels significantly increases in this condition, possibly indicting that the vesicle volume is larger (Fig. S6F, G). At 2dpf, the number of calcein stained pixels is similar in control and in continuous ROCK inhibition, suggesting that the overall calcium content does not change in this condition (Fig. S6H, I). The number of calcein stained vesicles is however decreased under ROCK inhibition, possibly indicating that there are fewer vesicles with larger volume. Addition of ROCK inhibitor at 25hpf affects neither the number of calcein stained vesicles nor the number of calcein stained pixels in the skeletogenic cells. Together, these measurements show that ROCK is not required for the uptake of calcium into the cells. Therefore, the skeletal phenotypes occurring under ROCK inhibition are not related to a decrease in calcium uptake or shortage in intracellular calcium, but are due to defects in mineral deposition.

To monitor the effect of ROCK inhibition on mineral deposition rate and distribution, we applied two pulses of different calcein dyes: blue calcein was added at 29hpf and washed after one hour, followed by green calcein added at 44hpf and washed after one hour. Hence, blue calcein marks the initial triradiate region and green calcein marks the edges of the spicules that were recently deposited (blue and green arrows in Fig. 3H, I). The dye FM4-64 was used to mark the cell membranes. Under ROCK inhibition, the green labeled region is much closer to the blue labeled region compared to control, due to the reduction in skeletal elongation rate, in agreement with our SR-µCT measurements. However, while in control embryos each body rod has a single tip stained in green calcein, under ROCK inhibition one of the body rods has two tips and both are stained in green. This indicates that mineral is being deposited in both edges, and the mechanism that prevents the growth of multiple tips in each rod, enabling “tip-dominance”, (in analogy to plant stem apical dominance), is disrupted under ROCK inhibition.

The activity of the actomyosin network is essential for normal spicule elongation and inhibition of ectopic branching

In other systems, major roles of ROCK are to control F-actin polymerization and MyoII activation (44–48), hence we wanted to directly test the role of these actomyosin components in sea urchin skeletogenesis. To directly inhibit F-actin polymerization we used Latrunculin-A (Lat-A), that prevents the polymerization of actin monomers (49) and was shown to be effective in sea urchin embryos (50). To directly inhibit actomyosin contractility we used Blebbistatin (Blebb), that interferes with MyoII ATPase activity, prevents actin-myosin interactions (51) and was shown to be effective in sea urchin embryos (52). To prevent interference with the early cell divisions where the actomyosin network plays a critical role (53–55), we added the inhibitors, individually or together, before or after spicule formation (20hpf and 25hpf, see methods for details). We tested the resulting phenotypes at 2dpf (Fig. 4A-G).

Our results indicate that F-actin polymerization and to a lesser extent, actomyosin contractility, are essential for normal skeletal growth and inhibition of ectopic branching. The majority of the embryos where MyoII activation was inhibited show mild ectopic branching mostly at the tips of the rods, and a minor delay in skeletal growth (Fig. 4B, C, G). The majority of the embryos where F-actin formation was inhibited show significant reduction of skeletal growth and severe ectopic skeletal branching (Fig. 4E, F and similar phenotypes to Fig. 4D, statistics in Fig. 4G). In some embryos, additional spicules form (Fig. 4F, G). The co-inhibition of F-actin polymerization and MyoII activity results in skeletogenic phenotypes that resemble, but are slightly stronger, than those of the inhibition of F-actin alone (Fig. 4D, G).

To test whether actin polymerization and myosin activity specifically in the skeletogenic cells underlies the observed skeletogenic phenotypes, we inhibit both in an isolated skeletogenic cell culture (Fig. 4H-K). We added the inhibitors after the spicule formed (>48hpf) and observed the phenotypes at 3dpf. Both inhibitors reduced skeletal growth, with Lantruculin-A having a significantly stronger effect than Blebbistatin (Fig. 4K), in agreement with the effect of these inhibitors in whole embryos. However, differently than in whole embryos, skeletal branching in both Lantruculin-A and Blebbistatin is not affected and is similar to the observed branching in control cultures. The absence of the branching phenotype in the skeletogenic cell culture could be due to the increased rigidity of the substrate that could compensate for the reduce actomyosin activity. Another option is that the branching phenotype in whole embryos could be due to the reduction of the activity of the actomyosin network in non-skeletogenic cells. Nevertheless, these results indicate that normal skeletal growth depends on F-actin polymerization and myosin contractility in the skeletogenic cells.

ROCK activity is required for F-actin organization around the forming spicule

Next, we tested whether ROCK regulates the actomyosin structure and activity during skeletal formation. To accomplish this we used phalloidin, a fluorophore that binds F-actin specifically, an antibody that detects phosphorylated MyoII (MyoIIP, (34)) and the skeletogenic marker, 6a9 (Fig. 5). In control embryos, F-actin is detected around the tri-radiate calcite spicules at 27hpf (green arrow in Fig. 5B, in agreement with (34, 35)). The pseudopodia cable that connects the skeletogenic cells shows lower F-actin signal compared to the signal around the spicule (blue line in Fig. 5B, C, E blue arrows in Fig. S7B, C, E). MyoIIP signal is not enriched in the skeletogenic cells nor in the spicule (Fig. 5D, S7D). Under continuous ROCK inhibition, the pseudopodia that connects the skeletogenic cells still forms (Fig. S7H) and F-actin accumulation is still observed in the lateral skeletogenic cell clusters; but unlike control embryos, F-actin is not organized and the spicule cavity is not formed (Fig. 5G). MyoIIP signal seems unchanged by ROCK inhibition (Fig. 5I, S7I), in agreement with the weak and late phenotype of MyoII inhibition (Figs. 4). These findings suggest that ROCK activity is essential for F-actin organization around the spicules, but not for the formation of the pseudopodia cable.

ROCK inhibition effect on F-actin organization and MyoII activity

A-T, Representative images showing normal embryos and embryos treated with 80µM ROCK inhibitor at different times. Right panels show enlargements of the rectangle sections marked in the left most panels. Phalloidin (green) marks F-actin, MyollP (red) marks phosphorylated MyoII and 6a9 (blue) marks the skeletogenic cells. A-E, Representative images at 27hpf showing control embryo and F-J, embryos treated with ROCK inhibitor from fertilization. Green arrows point to regions that show enriched F-actin signal. K-T, representative images at 33hpf, showing normal embryo K-O, and P-T, embryos where 30µM ROCK inhibitor was added at 25hpf. White arrowheads point to the enriched F-actin signal at the tips. Blue arrows point to the region of the pseudopodia cable that is not filled with the spicule cavity. The experiments were done in 3 biological replicates, the numbers at the bottom left of (A, F, K, P) indicate the number of embryos that show this phenotype out of all embryos scored. Scale bar in A, F, K, P is 50μm, and in B, G, L, Q is 10μm. U, V, representative spicules out of three biological replicates from control skeletogenic cell culture (n=60), U, and skeletogenic cell cultures that was treated with 30µM ROCK inhibitor was added at 48hpf (n=52), recorded at 72hpf. Left panel is phase image, middle panel is phalloidin staining and right panel shows the overlay. Green arrows point to the enhanced F-actin signal at the tips. Scale bar is 20μm.

F-actin is enriched at the tips of the elongating spicules independently of ROCK activity

To assess the role of ROCK in actomyosin organization during skeletal elongation we compared F-actin and myoIIP signals between control embryos and embryos where ROCK inhibitor was added at 25hpf (Fig. 5K-T, S7). At the prism stage, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Figs. 5L, O). We quantified the F-actin signal at the tips compared to the back in both treatments (Fig. S7K, L, see methods for details). The phalloidin signal is on average 3-fold stronger at the tips of the spicules compared to the back, in both control and ROCK inhibition, indicating that F-actin enrichment at the tips is significant and independent of ROCK activity. In both control and ROCK inhibited embryos, the F-actin signal was markedly higher in the regions of the pseudopodia cable where the spicule cavity had formed, compared to the regions where the spicule cavity had not yet formed (blue arrows in Fig. 5L-O, Q-T). We did not detect an enhanced signal of MyoIIP around the spicules or in the skeletogenic cells, in either condition (Fig. 5N, S). At the pluteus stage, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Fig. S7N-Q, S-V). Some non-skeletogenic cells that are enriched with MyoIIP signal are detected at this time (red arrowheads in Fig. S7N-P). Together these data demonstrate that F-actin coats the spicule cavity and is enriched at the tip of the rods, independently of ROCK activity.

We tested the effect of ROCK inhibition on F-actin in skeletogenic cell cultures. F-actin is detected at the tips, in both control cultures and cultures where ROCK was inhibited after spicule formation (Fig. 5U, V). Yet, under ROCK inhibition, branching is enhanced and F-actin is enriched at the splitting tips of the spicule rods (Fig, 5V). The enrichment of F-actin at the splitting tips of the spicule under ROCK inhibition in skeletogenic cell culture resembles the calcein staining at the two tips in the embryo in this condition (Fig. 3I). These observation further supports the role of ROCK activity in regulating tip-dominance during sea urchin skeletogenesis.

ROCK activity is essential for normal skeletogenic gene expression

The role of ROCK in regulating gene expression during vertebrates’ chondrocytes, osteoblasts and odontoblasts differentiation (20–22) intrigued us to study the role of ROCK in regulating skeletogenic gene expression in the sea urchin embryo. ROCK continuous inhibition significantly downregulates the expression of multiple skeletogenic genes, including the cytoskeleton remodeling gene, rhogap24l/2, the biomineralization genes, caral7, scl26a5 and SM30 at 27hpf and 2dpf (Fig. 6A, (6)). ROCK inhibition did not affect the expression level of VEGFR and ROCK itself at these times (Fig. 6A). Washing the inhibitor after 25hpf or adding the inhibitor at 25hpf had a weaker effect, but still led to downregulation of caral7, SM30, angio1 and notch1 (Fig. 6B). Thus, ROCK activity is required for the normal expression level of multiple skeletogenic genes during sea urchin skeletogenesis.

ROCK activity is essential for normal gene expression in the skeletogenic cells

A, B, the effect of 80µM ROCK inhibition on gene expression. A, continuous ROCK inhibition at 27hpf and 50 hpf (n=4). B, addition of ROCK inhibitor at 25hpf and the wash of ROCK inhibitor at 25hpf, measured at 50hpf (n=3). Asterisks indicate p<0.001, one tailed z-test. Error bars show standard deviation. C-F, Representative images of control embryo (top panels), embryos where ROCK inhibitor was added at 25hpf (middle panels), and embryos that were exposed to continuous ROCK inhibition (bottom panels), at the pluteus stage (∼48hpf). C, skeletogenic cell marker, 6a9. MV, midventral; AL, anterolateral, and PO, Post-oral rods, D, dorsal skeletogenic chain. D-F, spatial expression of skeletogenic genes. Gene names are indicated at the top of each panel. Numbers at the bottom of each image indicate the number of embryos that show this phenotype (left) out of all embryos scored (right), conducted in at least three independent biological replicates.

Both continuous and late ROCK inhibition, strongly affect the spatial expression of key regulatory and biomineralization genes that are normally expressed differentially within the skeletogenic lineage at the pluteus stage (Figs. 6D-F, S8). These include, the VEGF receptor (VEGFR), the biomineralization genes SM50 and SM30, skeletogenic transcription factors, Ets1, Alx1 and Erg1 that are essential for skeletogenic cell specification and biomineralization (23, 56) and the transcription factors Hex and MyoD1. In normal plutei these genes are differentially expressed within skeletogenic lineage, but in both continuous and late ROCK inhibition the expression of these genes becomes localized to the two lateral skeletogenic cell clusters and their vicinity (Fig. 6D-F, S8). The lateral skeletogenic cells clusters are the cells where these genes are expressed in the gastrula stage (24-27hpf) and where spicule formation begins in normal embryos (6, 32). The expression of VEGFR and SM50 under ROCK inhibition is more expanded than the other tested genes and is also observed at the anterolateral rods (Fig. 6B, C).The VEGF ligand that is normally expressed in the ectodermal cells near the four tips of the post-oral and anterolateral rods (32), is expressed at four ectodermal domains near the lateral skeletogenic cell clusters and near the end of the anterolateral chain, under ROCK inhibition (Fig. S8G).

Overall, under ROCK inhibition, the expression of key regulatory and biomineralization genes is downregulated (Fig. 6A) and remains localized to the area near the lateral skeletogenic cells clusters (Fig. 6D-F, S8). This localization is despite the proper formation of the dorsal, anterolateral and mid-ventral chains (Fig. 6C). This implies that ROCK activity provides regulatory cues to the skeletogenic GRN required for the dynamic progression of gene expression within the skeletogenic lineage.

Discussion

Biomineralization is a complex morphogenetic process, encoded in the species genome and executed by the GRNs and the cellular machinery. Here we used the sea urchin larval skeletogenesis to investigate the role of ROCK and the actomyosin network in biomineral formation and in regulating gene expression, as we summarize in Fig. 7. ROCK expression is enriched in the skeletogenic cells, partially depending on VEGF signaling (Fig. 1). ROCK activity in the skeletogenic cells is necessary for spicule formation, skeletal elongation and prevention of ectopic skeletal branching (Figs. 2-3, S3, 4). Direct inhibition of F-actin polymerization results in similar but not identical skeletogenic phenotypes (Fig. 4) and F-actin is organized around the spicule cavity and enriched at the spicule tips (Figs. 5, S7). ROCK activity feeds back into the skeletogenic GRN and affects regulatory and biomineralization gene expression (Fig. 6). Below we discuss our findings and their implications for the biological control and evolution of biomineralization.

The role of ROCK in sea urchin skeletogenesis, summary

A, at the gastrula stage (∼27hpf in P. lividus), the triradiate spicule is coated with F-actin. B, ROCK activity is required for spicule formation and F-actin organization. C, model of the regulatory interactions between the skeletogenic GRN and cytoskeleton remodeling during spicule initiation (∼24-27hpf in P. lividus). D, during skeletal elongation F-actin is detected around the spicule cavity and is enriched at the tips of the rods. The expression of SM50 and some other skeletogenic genes, is localized to the tips of the rods. E, under the addition of ROCK inhibitor, skeletal growth rate is reduced and ectopic spicule branching is observed. The expression of skeletogenic genes is localized to the vicinity of the growing rods. F, the regulatory interactions between the skeletogenic GRN and cytoskeleton remodeling during skeletal elongation.

The first step in generating the sea urchin spicules is the construction of the spicule cavity, a membrane filled with calcium carbonate and coated with F-actin (Fig. 7A). ROCK activity is necessary for the formation of the spicule cavity and for the organization of F-actin around it (Figs. 2, 5). A direct perturbation of F-actin polymerization strongly disrupts spicule growth and shape (Fig. 4). F-actin is enriched in the regions of the pseudopodia cable that are filled with mineral, supporting its role in mineral deposition (Figs. 5, S7, Fig. 7A). Overall, our studies indicate that ROCK activity in the skeletogenic cells is essential for the formation of the spicule cavity and for the initial F-actin organization around it, which could be essential for mineral deposition into it (Fig. 7C).

After the spicule cavity has formed, the spicules elongate inside the pseudopodia chord by mineral deposition at the tips of the spicule rods, which depends on ROCK activity and on F-actin polymerization (Figs. 2-4, S7, 7D-F). F-actin is detected along the spicules and is significantly enriched at the tips of the elongating spicules (Fig. 4, S7 (57)). F-actin enrichment at the tips is independent on ROCK activity (Fig. 5K-T, S7K, F), yet under ROCK inhibition and under the inhibition of F-actin polymerization, ectopic branching is observed at the tips and in the back (tip splitting and spicule dripping, Figs. 2J, N, 3A, 4, 5V). The participation of F-actin and ROCK in polarized growth and vesicle exocytosis has been observed in both animals and plants (57, 58). F-actin is used for short range myosin trafficking of vesicles toward the secretion site (57, 59). Once the secretory vesicle fuses with the membrane, it is coated with F-actin that stabilizes the vesicle and enables MyoII binding and ROCK-regulated contraction, necessary for content secretion (60, 61). In plant cells, F-actin accumulate at the growing tip of the cell and assists vesicle exocytosis necessary for polarized tip-growth (58). Our observations suggest that ROCK, F-actin and MyoII could be regulating the localized exocytosis of mineral-bearing vesicles during sea urchin skeletogenesis. The delay in spicule growth and the ectopic branching could be due to impaired and misplaced vesicle trafficking and deposition under ROCK and F-actin inhibition, but the exact protein regulatory network is still to be determined.

ROCK activity is necessary for correct spatiotemporal expression of regulatory and biomineralization genes (Fig. 6). ROCK inhibition downregulates the expression level of skeletogenic genes (Fig. 6A) and prevents the spatial progression of key skeletogenic genes toward the edges of the skeletal rods (Fig. 7E). In other systems, ROCK activity was shown to regulate gene expression through various intracellular pathways (62–64). Furthermore, ROCK has an important role in mechano-sensing of matrix rigidity and transducing it into changes in gene expression (65). Relatedly, stiff matrix activates ROCK in pre-osteoblastic cells, which activates Erk that leads to osteogenic differentiation and upregulation of osteogenic gene expression (64). Sea urchin ROCK could be sensing the spicule stiffness and transducing it into gene expression, providing a necessary feedback between spicule elongation and the skeletogenic GRN. Further studies are required to identify the specific molecular mechanisms activated by ROCK to affect gene expression in the skeletogenic cells.

Overall, our findings and the role of actomyosin remodeling in biomineralization, from calcifying and silicifying single cells to vertebrates’ bones and teeth (15-22, 64), suggest that this molecular machinery is a part of the common molecular tool-kit of biomineralization. Most likely, the actomyosin network was employed independently, downstream of distinct GRNs across Eukaryotes, to control mineral growth and morphogenesis.

Methods

Animal and embryos

Adult Paracentrotous lividus were obtained from the Institute of Oceanographic and Limnological Research (IOLR) in Eilat, Israel. Eggs and sperm were obtained by injection of 0.5M KCl solution into the adult sea urchin species. Embryos were cultured at 18°C in 0.2µ filtered ASW.

Imaging

Embryonic phenotypes and WMISH were imaged by Zeiss Axioimager 2. Fluorescent markers were imaged using Nikon A1-R Confocal microscope. All images were aligned in Adobe photoshop and Illustrator.

Axitinib (AG013736) treatment

Axitinib (AG013736, Selleckchem, Houston, TX, USA) was applied as described in (32)

Western Blot

Embryo cultures treated with 150 nM Axitinib, or DMSO as control, were collected at 22hpf, 27hpf and 33hpf by centrifugation of 10,000g for 2 minutes. We lysed the pellet in 150µL lysis buffer (20 mM Tris-HCl, 150 mM NaCl, 1% Triton X-100, pH 8) with protease inhibitors (protease inhibitor cocktail w/o metal chelator; Sigma, P8340; 1:100) and phosphatase inhibitors (50 mM NaF, 1 mM Na3VO4, 1 mM Na4P2O7, 1mM, βGP) as described in (66). 40µg protein were loaded on 8% SDS-acrylamide gel, transferred to PVDF membranes for Western blot analysis. Anti-ROCK antibody (ab45171, Abcam) was used in a 1:300 dilution followed by anti-rabbit HRP secondary antibody diluted to 1:2500 (Jackson ImmunoResearch 111-035-003). For the loading control, the membrane was incubated with anti-β-tubulin antibody (1:5000; Sigma) followed by anti-mouse antibody 1:5000 (Jackson ImmunoResearch 115-035-003). Quantification of ROCK signal was done using Image studio lite vr. 5.2. ROCK signal was divided by Tubulin signal and then the ratio between ROCK/Tubulin in control vs. VEGFR inhibition was calculated. The graph in Fig. S2 was generated based on three biological replicates (different sets of parents) at 27hpf and 33hpdf and four biological replicates at 22hpf.

Immunostaining procedure

Phalloidin labeling, MyoIIP (p20-MyosinII), ROCK (anti ROCK2+ROCK1 antibody [EP786Y], AB-ab45171, Abcam), and 6a9 immunostaining were done similarly to (34).

Quantification of Anti-ROCK and Phalloidin signal

A graphical program was written to allow manual quantification of the fluorescent signal from the spicule regions by identifying “stained” pixels per selected area. The ratio of stained anti-ROCK to region total area was compared between ectodermal and skeletogenic cells, and between control and VEGFR inhibited embryos at 22, 27 and 33hpf. The average ratios and ztest for significance difference from 1, are 22hpf control, ∼1.3, z=0.0006; VEGFR inhibition ∼1, z=0.5, 27hpf control ∼1.8, z<10-11; VEGFR inhibition ∼1.6, z<10-8, 33hpf control ∼4.6, z<10-23; VEGFR inhibition ∼1.5, z<10-5. For phalloidin analyses, embryos at 33hpf of control and ROCK inhibitor addition at 25hpf were used and for each image, two regions were selected: an area at the tip of the spicule and an area at the back of the spicule, about ∼10 microns apart. The ratio of stained phalloidin to region total area was compared between tip and back, in control and ROCK inhibited embryos. Paired t-test was used to compare the differences in phalloidin signal between the tip and back of the spicules across all groups and between control and ROCK inhibited embryos.

ROCK MASOs injections

Microinjections were conducted similarly to (32). The eggs were injected with an injection solution containing 0.12M KCl, 0.5µg/µl Rhodamine Dextran and 900µM of MASOs. Translation MASOs were designed and synthesized by Gene Tools, Philomath, OR, according to ROCK sequence (39). The sequence of ROCK MASO-1 (5’-AGACATATTTGGAGCCGA[CAT]CCTG-3’) matches the start of translation region and ROCK MASO-2 (5’-TCTCTTGCGTTATATTCCACTAAGT-3’) matches the 5’ region upstream of the start of translation (39). Control embryos were injected in the same concentration with Random MASO. Injected embryos were cultured at 18°C, imaged and scored at 2dpf.

ROCK inhibitor (Y27632) treatment

Y27632 stock (10005583 CAS Regisrty No. 129830-38-2, Cayman chemical), was prepared by reconstituting the chemical in DMSO. The embryos were treated with Y27632 at a final concentration between 30-80µM, as mentioned in the results. Throughout the paper, control embryos were cultured in DMSO at the same volume as Y27632 solution and no more than 0.1% (v/v).

ROCK inhibition in isolated skeletogenic cell culture

Skeletogenic cell culture was performed as described in (36) with minor changes. Isolated skeletogenic cells were cultured in CultureSlides (4 chambers polystyrene vessel tissue culture treated glass slide REF 354104) with ASW (Red Sea Fish LTD) containing Gentamicin and Penicillin-Streptomycin (GPS) at 18°C. Media containing 4% (v/v) horse serum (Sigma-Aldrich H1270) in ASW+GPS was added to the culture. ROCK inhibitor, Y27632, was added to the cell culture at 16hpf (before spicule initiation) or 48hpf (after spicule initiation), and images were taken at 72hpf. Each treatment was conducted in three biological repeats except from the >48hpf addition of 120µM which was done in two biological repeats.

Quantification of skeletal length and statistical analysis

Skeletal length measurement was done as described in (32). The measurements were repeated for three biological repeats for control, ROCK inhibition with 30μM and 80μM, Latrunculin-A 2nM, Blebbistatin 2µM, and 2 biological repeats for 120μM. In ROCK inhibition experiments, a total of 275 skeletons were measured for control, 105 skeletons for 30μM, 122 skeletons for 80μM and 80 for 120μM. In Latrunculin-A and Blebbistatin experiments, a total of 116 skeletons were measured for control, 149 for Latrunculin-A and 107 for Blebbistatin. The data were analyzed in Excel and the statistical analysis was performed using Kruskal-Wallis non-parametric test in Excel as described in (67) and SPSS statistics 27.

Spicule preparation for micro-computed tomography (SR-µCT) measurement

Control embryos and embryos where 40µM Y27632 was added at 25hpf, were collected by centrifugation of 50mL Falcon tubes at 2000rpm, at 48h and 72hpf. Every tube was washed in 10 mL cold distilled water (DW), transferred into epi-tube and washed three times in 1mL cold DW (incubation at 4°C between every wash, centrifuge at 2000rpm in room temperature for 2min). Then the skeletons were washed three times in 3% NaOCl followed by three washes in DW, one wash in 70% EtOH and one wash in Acetone. Skeletal spicules were dried at room temperature and stored at −20 °C.

Synchrotron and lab-based SR-µCT

Dry and loose skeletal spicules prepared as described above, were used for the SR-µCT analysis. We analyzed 14 sets of spicules from three (control) and four (ROCK inhibited) pairs of parents, divided into four groups: 2dpf, control, n=44; 2dpf ROCK inhibited, n=93; 3dpf control, n=51; 3dpf ROCK inhibited, n=93, see Table S1). Spicules were glued to sharpened toothpick tips, either in bulk or individually when picked by hand by means of a brush bristle. Each toothpick loaded with spicules, was mounted on a 3ml vial twist-off lid and for transportation securely stored in the vial (Fig. S5A, B). For each set 2-3 µCT-samples were prepared, which allowed the acquisition of statistically relevant sample sizes (40+ spicules/group, Table S1). Tomographic data sets were acquired using the following scanning parameters; for the lab based µCT (RX Solutions, Chavanod, France) 90kV, 143-160µA, 1440 images, average frames 10, frame rates 0.4-1, voxel sizes: 0.45-59µm; and for the synchrotron radiation µCT (Anatomix, Synchrotron SOLEIL, Gif-sur-Yvette, France) 2048 images, angle between projections 0.09°, voxel size: 0.65µm. Tomographic data reconstruction was performed with commercial (RX Solutions, Chavanod, France) and PyHST2 software (A. Mirone et al., Nucl. Instrum. Meth. B 324 (2014) 41-48, doi:10.1016/j.nimb.2013.09.030) at the MPICI and at Anatomix, SOLEIL, respectively.

Three-dimensional (3D) data analysis

Data visualization, pre-processing including cropping and image type conversion to 16bit (of the Sr-µCT data), and user-augmented segmentation of intact spicules were performed in Amira using off-the-shelf modules (Fig. S5C-E, (68)). Volume and area measurements were performed on the segmented spicules, i.e. the label-fields using Amira’s material.statistics module (Volume3D and Area3D). Average spicule thickness and total length measurements were performed with python code, utilizing available python libraries (e.g., skimage, skan, scipy). The code was applied to 3D-tif files created from Amira’s label-fields, which contained the segmented spicules. For length measurements, the spicules were skeletonized and the chord lengths of the branches forming a spicule-skeleton were summed up to obtain the total spicule length. For thickness measurements, we performed Euclidean distance transforms for each spicule using scipy’s ndimage.morphology.distance_transform_edt package, and calculated the average spicule thickness based on the distance measurements obtained at each voxel from spicule-skeleton to the nearest spicule surface. Quality control of our measurements was performed with open source software, Fiji and available plugins (e.g. BoneJ, FiberJ, DiameterJ).

Calcein staining

A 2mg/ml stock solution of blue calcein (M1255, Sigma) and a 2mg/ml stock solution of green calcein (C0875, Sigma) were prepared by dissolving the chemicals in distilled water. Working solution of 250μg/ml and 150μg/ml, respectively, was prepared by diluting the stock solution in artificial sea water. Blue calcein was added to the embryo culture at 29hpf for one hour and then washed. At 44hpf the green calcein was added for one hour and washed. For FM4-64 staining A 100µg/ml stock solution of FM4-64 (T13320, Life technologies, OR, USA) was prepared by dissolving the chemical in distilled water. Working solution of 2.5 µg/ml was prepared by diluting the stock solution in artificial sea water. Embryos were immersed in working solution about 10 minutes before visualization.

Latrunculin-A and Blebbistatin treatments

1mM of Lat-A stock (L12370, Thermo Fisher Scientific) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Lat-A at final concentration of 2 nM. 1mM of Blebbistatin stock (13013-5, Enco) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Blebbistatin at final concentration of 2 µM. For the double inhibition experiments, embryos were treated with Blebb at final concentration of 1.5 µM, and with Lat-A at final concentration of 2 nM.

WMISH procedure

WMISH was performed as described in (32). List of the primers used to make the WMISH probes is available in (32).

cDNA preparation for QPCR experiments

For ROCK inhibition experiments total RNA was extracted from >1000 sea urchin embryos in each condition using RNeasy Mini Kit (50) from QIAGEN (#74104) according to the kits’ protocol. DNase treatment on column was done using RNase-Free DNase Set-Qiagen (50) (#79254). RNA was reverse-transcribed using the High Capacity cDNA RT kit, AB-4368814 (Applied Biosystems) following the manufacturer’s instructions.

Quantitative polymerase chain reaction (QPCR)

QPCR was carried out as described in (32). Complete list of primer sequences used for the experiments in provided there.

Supporting information

eLife Assessment

This valuable paper addresses a notable problem, the cell biological control of biomineralization, with the sea urchin embryo as an experimental model. The paper provides evidence that ROCK and the cytoskeleton play a role in biomineralization, but the evidence is deemed currently incomplete, as there are concerns regarding the efficacy and specificity of the reagents used to perturb ROCK function. In addition, the data do not point to a plausible mechanism by which the actin cytoskeleton might regulate the biomineralization process.

Reviewer #1 (Public Review):

Using a pharmacological and knock-down approach, the authors could demonstrate that ROCK activity is required for the normal development of the larval skeleton. The presence of ROCK in the pluteus stage depends on the activity of VEGF that is responsible for the formation of the tubular syncytial sheath of the calcifying primary mesenchyme cells in which the skeleton forms. The importance of ROCK in skeleton formation was confirmed in cell culture experiments, demonstrating that ROCK inhibition leads to decreased elongation and abnormal branching of spicules. µCT analyses underline this finding demonstrating that the inhibition of ROCK mainly affects the elongation of spicules while growth in girth is little affected. F-actin labeling experiments could demonstrate that ROCK inhibition interferes with the organization of the actomyosin network in the early phase of skeleton formation, while f-actin organization in the tips of the elongating spicule is unaffected by the pharmacological inhibition of ROCK. Finally, ROCK inhibition strongly affects the expression of major regulatory and calcification-related genes in the calcifying cells. Based on these findings the authors propose a model for the regulatory interaction between the skeletogenic GRN, ROCK, and the f-actin system relevant for skeletogenesis.

I reviewed this paper previously for submission to another Journal. I emphasize again, that this is an interesting and important work that aims to uncover the interaction between the Rho-associated Kinase, ROCK, the actomyosin network, and its relevance for the formation of the calcitic skeleton of the sea urchin larva. I carefully went through the revised manuscript. In their new version, the authors rearranged the figures to provide a more direct comparison between the in vivo and cell culture experiments which mitigates the criticism of collateral effects by the inhibitors on the whole organism. The authors also performed an additional experiment localizing the F-Actin signal in spicules of PMC cell cultures under ROCK inhibition. This experiment strengthens the concept that ROCK activity is important for tip dominance rather than CaCO3 deposition rates. The results section was substantially reorganized and only very minor changes were made to the introduction and discussion.

I think that this work has great potential to provide seminal insights into an understudied aspect of the biomineralization process - the role and regulation of the cytoskeleton in calcifying cells. As I mentioned in my previous review there are some gaps in this work that need to be answered to provide a conclusive dataset on the role of ROCK and the actomyosin system in the mineralization process. The manuscript in its current form provides evidence for the interaction of ROCK with the actomyosin system in the sea urchin larva and that perturbation of this system affects skeletogenesis. However, it is missing an explanation regarding the mechanism by which ROCK affects skeleton formation. No difference in f-actin localization was found at the spicule tips in control and ROCK-inhibited larvae. A slight hint was found in the difference in vesicle size and f-actin organization within calcifying cells, but it remains unresolved if ROCK activity impacts the trafficking of calcification vesicles. The authors provide an interesting discussion on the involvement of f-actin and ROCK on vesicular trafficking, and exocytosis based on existing knowledge from animal and plant models. But for the sea urchin larva, this important link between ROCK, f-actin, and the biomineralization process remains unanswered. In their previous work by Winter et al. 2021, the authors demonstrated excellent technologies to monitor vesicular dynamics in the calcifying cells. This tool would be ideal to investigate the role of ROCK and the actomyosin network on the trafficking dynamics of Ca2+-rich vesicles. These experiments (among others suggested in the following review) may help to uncover the critical mechanism to resolve the missing gap in this manuscript.

Major comments

One MASO led to reduced skeleton formation while the other one additionally induced ectopic branching. How was the optimum concentration for the MASOs determined? Did the authors perform a dose-response curve? What is the reason for this difference? Which of the two MASOs can be validated by reduced ROCK protein abundance? Since the ROCK antibody works, I would like to see a control experiment on Rock protein abundance in control and ROCK MO injected larvae which is the gold-standard for validating the knock-down.

L212 "Together, these measurements show that ROCK is not required for the uptake of calcium into cells." But what about trafficking and exocytosis? As mentioned earlier, I think this is a really important point that needs to be confirmed to understand the function of ROCK in controlling calcification. In their previous study (reference 45) the authors demonstrated that they have superior techniques in measuring vesicle dynamics in vivo. Here an acute treatment with the ROCK inhibitor would be sufficient to test if calcein-positive vesicle motion, including the observed reduction in velocity close to the tissue skeleton interface, is affected by the inhibitor.

Is there a colocalization of ROCK and f-actin in the tips of the spicules? This would support the mechano-sensing-hypothesis by ROCK.

L 283. "F-actin is enriched at the tips of the spicules independently of ROCK activity" The results of this paragraph clearly demonstrate that ROCK inhibition has no effect on the localization of f-actin at the tips of the growing spicules. In addition, the new cell culture experiments underline this observation. Still, the central question that remains is, what is the interaction between ROCK, f-actin, and the mineralization process, that leads to the observed deformations? What does the f-actin signal look like in a branched phenotype or in larvae that failed to develop a skeleton (inhibition from Y20)?

Immunohistochemical analyses on f-actin localization and abundance should be additionally performed with ROCK knock-down phenotypes to confirm the pharmacological inhibition.

L 365 "...supporting its role in mineral deposition..." "...Overall, our studies indicate that ROCK activity....is essential for the formation of the spicule cavity......which could be essential for mineral deposition..." I think the authors need to do a better job in clearly separating between the potential processes impacted by ROCK perturbation. Is it stabilization and mechano-sensing in the spicule tip or the intracellular trafficking and deposition of the ACC? If the dataset does not allow for a definite conclusion, I suggest clearly separating the different possibilities combined with thorough discussion-based findings from other mineralizing systems where the interaction between ROCK and F-actin has been described.

Reviewer #2 (Public Review):

This manuscript reports on the role of Rho-associated coiled-coil kinase (ROCK) in biomineralization of sea urchin larval skeletons. A number of experiments examine the initiation, growth, and patterning of the skeleton in an effort to determine if, and how, ROCK participates in skeletal formation. The authors conclude that ROCK controls the formation, growth, and morphology (patterning) of the skeleton based on a number of inhibition studies. The main target of the experiments is the actomyosin cytoskeleton which has been the focus of many ROCK studies in vertebrates. Based on similar experimental outcomes when comparing the results here with published data from vertebrates, they suggest that ROCK and the actomyosin network operate in a similar way in biomineralization despite independent evolutionary origins of the sea urchin larval skeletons and the skeletons of vertebrates.

My concerns are the interpretation of the experiments. The main overriding concern is a possible over-interpretation of the role of ROCK. In the literature that ROCK participates in many biological processes with a major contribution to the actin cytoskeleton. And when a function is attributed to ROCK, it is usually based on the determination of a protein that is phosphorylated by this kinase. Here that is not the case. The observation here is in most cases stunted growth of the spicule skeleton and some mis-patterning occurs or there is an absence of skeleton if the inhibitor is added prior to initiation of skeletal growth. They state in the abstract that ROCK impairs the organization of F-actin around the spicules. The evidence for that as a direct role is absent. They use morpholino data and ROCK inhibitor data to draw their conclusion. My main concern is the concentration of the inhibitor used since at the high concentrations used, the inhibitor chosen is known to inhibit other kinases as well as ROCK (PKA and PKC). They indicate that this inhibition is specifically in the skeletogenic cells based on the isolation of skeletogenic cells in culture and spicule production either under control or ROCK inhibition and they observe the same - stunting and branching or absence of skeletons if treated before skeletogenesis commences. Again, however, the high concentrations are known to inhibit the other kinases. They use blebbistatin and latrunculin and show that these known inhibitors of actin cytoskeleton lead to abnormal spiculogenesis, This coincidence is suggestive but is not proof that it is ROCK acts on the actomyosin cytoskeleton given the specificity concerns.

Author Response

The following is the authors’ response to the original reviews.

Thank you and the two reviewers for the thorough review of our manuscript. We found the reviewer’s comments highly valuable and addressed them by the following additional experiments and changes in the text and the figures:

(1) We measured the effect of ROCK MASO’s on the ROCK expression by immunostaining and observed a reduction in ROCK signal, supporting the downregulation of ROCK protein level under ROCK MASO’s (new Fig. S3).

(2) We measured the effect of lower concertation of ROCK inhibitor, Y27632 (10µM), and observe the same phenotypes of skeletal loss, skeletal reduction and ectopic branching in this concentration (Fig. 2, S4). Importantly, these phenotypes were not observed when directly inhibiting PKA and PKC, in whole sea urchin embryos (1) and in skeletogenic cell cultures (2), further supporting the specificity of ROCK inhibitor.

(3) We added a time course of Pl-ROCK expression and immunostaining of ROCK in the fertilized egg, that show that this gene is maternal and the protein is present in the egg Fig. 2SA-C.

(4) We recorded F-actin in ROCK MASO’s and demonstrate that it is still detected around the spicules and their tips, similarly to ROCK inhibited embryos (new Fig.S3).

(5) We revised the paper text and figures to provide a better description of our results, distinguish clearly between our data and our interpretations and emphasize the novelty of our findings.

This paper demonstrates that ROCK, F-actin polymerization and actomyosin contractility play critical roles in biomineral growth and in shaping biomineral morphology in the sea urchin embryo, and that ROCK activity affects skeletogenic gene expression. Our findings together with previous reports of the role of actomyosin in Eukaryotes biomineralization, suggest that this molecular machinery is a part of the common molecular tool-kit used in biomineralization. The identification of a common molecular mechanism within the diverse gene regulatory networks, organic scaffolds and minerals that Eukaryote use to build their biominerals will be of high interest to the field of biomineralization and evolutionary biology. Furthermore, our paper portrays the interplay between the cellular and the genetic machinery that drives morphogenesis. We believe it would be of great interest to the broad readership of eLife and particularly to the fields of biomineralization, cell, developmental and evolutionary biology.

Thank you very much for the helpful review of our paper.

Reviewer #1 (Public Review):

We thank the reviewer for the appreciation of our work the helpful comments that guided us to strengthen the experimental evidence for our conclusions and increase the paper’s clarity. Below are our responses to the specific comments:

Major comments

One MASO led to reduced skeleton formation while the other one additionally induced ectopic branching. How was the optimum concentration for the MASOs determined? Did the authors perform a dose-response curve? What is the reason for this difference? Which of the two MASOs can be validated by reduced ROCK protein abundance? Since the ROCK antibody works, I would like to see a control experiment on Rock protein abundance in control and ROCK MO injected larvae which is the gold-standard for validating the knock-down.

We tested several MASO concentrations to identify a concentration where the control embryos injected with Random MASO were overall healthy and ROCK MASO’s showed clear phenotypes.

To test the effect of ROCK MASO’s on ROCK protein levels we did immunostaining experiments that are now presented in new Fig. S3. We could not do Western blot for injected embryos since ROCK antibody requires thousands of embryos for Western blot, which is not feasible for injected embryos. Therefore, we tested the effect of the two translation ROCK MASO’s on ROCK abundance compared to uninjected and Random MASO injected embryos using immunostaining. We observed a reduction of ROCK signal, supporting the downregulation of ROCK protein level in these genetic perturbations (new Fig. S3).

L212 "Together, these measurements show that ROCK is not required for the uptake of calcium into cells." But what about trafficking and exocytosis? As mentioned earlier, I think this is a really important point that needs to be confirmed to understand the function of ROCK in controlling calcification. In their previous study (reference 45) the authors demonstrated that they have superior techniques in measuring vesicle dynamics in vivo. Here an acute treatment with the ROCK inhibitor would be sufficient to test if calcein-positive vesicle motion, including the observed reduction in velocity close to the tissue skeleton interface, is affected by the inhibitor.

We thank the reviewer for the appreciation of our previous work where we studied calcium vesicle dynamics in whole embryos (Winter et al, Plos Com Biol 2021). We agree with the reviewer that the best way to test directly the effect of ROCK on mineral deposition and vesicle kinetics is to observe it in live skeletogenic cells. However, in Winter et al 2021, we found that the skeleton (spicules) doesn’t grow when the embryos are immobilized in either control or treated embryos. We have to immobilize the embryos to record live timelapses of whole embryos. Hence, this means that we can not determine the role of ROCK or any other perturbation in vesicle trafficking and exocytosis based on experiments conducted in immobilized whole embryos, since skeletogenesis is arrested. We believe that we can do it in skeletogenic cell cultures and we are currently developing this assay for vesicle tracking, but this is beyond the scope of this current work.

Is there a colocalization of ROCK and f-actin in the tips of the spicules? This would support the mechano-sensing-hypothesis by ROCK.

Our studies show that F-actin is localized around the spicule cavity and in the cortex of the cells (Figs. 5 and 6) while ROCK is enriched in the skeletogenic cell bodies, with some localization near the skeletogenic cell membranes (Fig. 1). To directly address the reviewer question we immune-stained ROCK and F-actin in the same embryos, and showed that their sub-cellular localizations does not show a strong overlap (Fig. S3 Q-T). However, ROCK does not bind F-actin directly: ROCK activates another kinase, LimK that phosphorylates Cofilin that interacts with F-actin. Therefore, the fact that ROCK is not colocalized with F-actin does not support nor contradicts the possible role of ROCK in mechano-sensing.

L 283. "F-actin is enriched at the tips of the spicules independently of ROCK activity" The results of this paragraph clearly demonstrate that ROCK inhibition has no effect on the localization of f-actin at the tips of the growing spicules. In addition, the new cell culture experiments underline this observation. Still, the central question that remains is, what is the interaction between ROCK, f-actin, and the mineralization process, that leads to the observed deformations? What does the f-actin signal look like in a branched phenotype or in larvae that failed to develop a skeleton (inhibition from Y20)?

As we report in Fig. 6, and now on new Fig. S3, under ROCK late inhibition or in ROCK morphants, we still detect F-actin around the spicule and enriched at the tips. When ROCK is inhibited and the embryo fails to develop a skeleton, we observe Factin accumulation in the skeletogenic cells, but the F-actin is not organized (Fig. 5). As the spicule is absent in this condition, it is hard to conclude whether the effect on F-actin organization is direct or due to the absence of spicule in this condition. We stated that explicitly in the current version in the results, lines 324-326 and in the discussion, lines 405-408.

Immunohistochemical analyses on f-actin localization and abundance should be additionally performed with ROCK knock-down phenotypes to confirm the pharmacological inhibition.

We did that in our new Figure S3 and showed that ROCK morphant show the same F-actin localization at the tips like control and ROCK inhibited embryos.

L 365 "...supporting its role in mineral deposition..." "...Overall, our studies indicate that ROCK activity....is essential for the formation of the spicule cavity......which could be essential for mineral deposition..." I think the authors need to do a better job in clearly separating between the potential processes impacted by ROCK perturbation. Is it stabilization and mechano-sensing in the spicule tip or the intracellular trafficking and deposition of the ACC? If the dataset does not allow for a definite conclusion, I suggest clearly separating the different possibilities combined with thorough discussion-based findings from other mineralizing systems where the interaction between ROCK and F-actin has been described.

We thank the reviewer for this important comment. We believe that ROCK and the actomyosin are involved in both, mechano-sensing of the rigid biomineral and in the transport and exocytosis of mineral-bearing vesicles. In the current version we provide explicit explanations of these two hypotheses in the discussion section. The possible role in exocytosis and the experiments that are required to assess this role are described in lines 427-439, and the possible mechano-sensing role and effect on gene expression is described in lines 440-453.

Reviewer #1 (Recommendations For The Authors):

Minor comments

L185 "These SR-µCT measurements show that the rate of mineral deposition is significantly reduced under ROCK inhibition." To correctly support this statement I would suggest to calculate the real growth rates (µm3 time-1). For example, an increase in volume from 6,850 µm3 at 48 hpf to 14,673 µm3 at 72 hpf would result in a growth rate of 7823 µm3 24h-1.

We thank the reviewer for this suggestion. We calculated the rate of spicule growth as the reviewer suggested and we added this information in lines 218-221.

L343: "This implies that....within the skeletogenic lineage." This concluding sentence is very speculative and therefore misplaced in the results section.

We removed this sentence from the results section into the discussion, lines 443-445.

L382: "The participation of F-actin and ROCK in polarized tip-growth and vesicle exocytosis has been observed in both, animals and plants." L407-409: "...F-actin could be regulating the localized exocytosis of mineral-bearing vesicles...." I think this is exactly the core question that remains unresolved in this study. To reduce speculations I strongly recommend addressing the effect of ROCK inhibition on vesicle trafficking and exocytosis (Monitoring of calcein-positive Vesicles in PMCs).

We agree with the reviewer that this is a critical question that we would have address, but as we explained above, is beyond the scope of this study.

Figure 5: The values below the scale bars in the newly added figures U+V are extremely small. Also, the Legend for this figure sounds incorrect. Should read: "...and skeletogenic cell cultures that were treated with 30µM ROCK inhibitor that was added at 48hpf and recorded at 72hpf.

We increased the font near the scale bars and corrected the figure caption. Thanks for this and your other helpful comments!

Reviewer #2 (Public Review):

We thank the reviewer for raising the important issue of inhibitor concentration which led us to do additional experiments with lower concentration that were valuable and strengthen the manuscript. We also thank the reviewer for asking us to be clearer with the interpretation of the results. Below are our responses to the specific comments:

My concerns are the interpretation of the experiments. The main overriding concern is a possible over-interpretation of the role of ROCK. In the literature that ROCK participates in many biological processes with a major contribution to the actin cytoskeleton. And when a function is attributed to ROCK, it is usually based on the determination of a protein that is phosphorylated by this kinase. Here that is not the case. The observation here is in most cases stunted growth of the spicule skeleton and some mis-patterning occurs or there is an absence of skeleton if the inhibitor is added prior to initiation of skeletal growth. They state in the abstract that ROCK impairs the organization of F-actin around the spicules. The evidence for that as a direct role is absent.

We agree with the reviewer that since the spicule doesn’t form under ROCK continuous inhibition, it is unclear if the absence of F-actin around the spicule in this condition is a direct outcome of the lack of ROCK activation of F-actin polymerization, or an indirect outcome due to the lack of spicule to coat. We therefore deleted this line in the abstract and explicitly stated that we cannot conclude whether the impaired F-actin organization is directly due to ROCK effect on actin polymerization in the results, lines 324-326 and in the discussion, lines 405-408.

They use morpholino data and ROCK inhibitor data to draw their conclusion. My main concern is the concentration of the inhibitor used since at the high concentrations used, the inhibitor chosen is known to inhibit other kinases as well as ROCK (PKA and PKC). They indicate that this inhibition is specifically in the skeletogenic cells based on the isolation of skeletogenic cells in culture and spicule production either under control or ROCK inhibition and they observe the same - stunting and branching or absence of skeletons if treated before skeletogenesis commences. Again, however, the high concentrations are known to inhibit the other kinases.

In the previous version of the paper we used the range of 30-80µM Y-27632 to block ROCK activity. These concentrations are commonly used in mammalian systems and in Drosophila to block ROCK activity (3-8). The reviewer is correct stating that at high concentration, this inhibitor can block PKA and PKC. However, the affinity of the inhibitor for these kinases is more than 100 times lower than its affinity to ROCK as indicated by the biochemical Ki values reported in the manufactory datasheet: 0.14-0.22 μM for ROCK1, 0.3 μM for ROCK2, 25 μM for PKA and 26 μM for PKC.

Importantly, these Ki values are based on biochemistry assays where the activity of the inhibitor is tested in-vitro with the purified protein. Therefore, these concentrations are not relevant to cell or embryo cultures where the inhibitor has to penetrate the cells and affect ROCK activity in-vivo. Y-27632 activity was studied both in-vitro and in-vivo in Narumiya, Ishizaki and Ufhata, Methods in Enzymology 2000 (9). This paper reports similar concentrations to the ones indicated in the manufactory datasheet for the in-vitro experiments, but shows that 10µM concentration or higher are effective in cell cultures. We therefore tested the effect of 10µM Y-27632 added at 0hpf (continuous inhibition) and at 25hpf (late inhibition) and added this information to Figs. 2 and S3. Continuous inhibition at this concentration resulted with three major phenotypes: skeletal loss, spicule initiations and small spicules with ectopic branching. This result supports our conclusion that ROCK activity is necessary for spicule formation, elongation and prevention of branching. Late inhibition in this concentration resulted with the majority of the embryos developing branched spicules, which is very similar to the effect of MyoII inhibition with Blebbistatin. This result again, supports the inference that ROCK activity is required for normal skeletal growth and the prevention of ectopic branching. Importantly, there are two papers were PKA and PKC were directly inhibited in whole sea urchin embryos (1) and in skeletogenic cell cultures (2). In both assays, PKC inhibition resulted with mild reduction of spicule length while PKA inhibition did not affect skeletal formation. Neither skeletal loss nor ectopic branching were ever observed under PKC or PKA inhibition, supporting the specific inhibition of ROCK by Y-27362. Furthermore, both genetic and pharmacological perturbations of ROCK resulted with significant reduction of skeletal growth and with the enhancement of ectopic branching. Therefore, we believe we provide convincing evidence for the role of ROCK in spicule formation, growth and prevention of branching. We revised Fig. 2 and S3 to include the 10µM Y-27632 data and the text describing the inhibition to include the explanations and references we provided here.

They use blebbistatin and latrunculin and show that these known inhibitors of actin cytoskeleton lead to abnormal spiculogenesis, This coincidence is suggestive but is not proof that it is ROCK acts on the actomyosin cytoskeleton given the specificity concerns.

As stated above, we believe that in the current vesion we overcame the specificity concerns and provided solid evidence that ROCK activity is necessary for spicule formation, growth and prevention of branching. Furthermore, the skeletogenic phenotypes of late 10µM Y-27632 are highly similar to those of MyoII inhibition (Blebbistatin) while the phenotypes of higher concetrations resemble the inhibition of actin polymerization by Latrunculin. We agree with the reviewer that: “This coincidence is suggestive but is not proof that ROCK acts on the actomyosin cytoskeleton” and we revise the discussion paragraph to differentiate between our solid findings and our speculations (lines 421-426): “These correlative similarities between ROCK and the actomyosin perturbations lead us to the following speculations: the low dosage of late ROCK inhibition is perturbing mostly ROCK activation of MyoII contractility while the higher dosage affects factors that control actin polymerization (Fig. 8F). Further studies in higher temporal and spatial resolution of MyoIIP activity and F-actin structures in control and under ROCK inhibition will enable us to test this.”

Reviewer #2 (Recommendations For The Authors):

The following areas require attention:

(1) You begin and end the abstract with statements on evolution in which the actomyosin cytoskeleton is associated with skeletogenesis despite different GRNs, different contributing proteins, etc. You then move to ROCK and claim to reveal that ROCK is a central player in the process. As above, in the judgement of this reviewer, you fail to establish a direct role of ROCK to the actomyosin role in skeletogenesis. Sure, the ROCK inhibitors suggest that ROCK plays some kind of role in the process but you also indicate that ROCK could act on many processes, none of which you directly associate with the necessary activity of ROCK.

We agree that our paper provides correlative similarities between the phenotypes of ROCK and those of direct pertrubations of the actomyosin network, and lacks causal relationship. We made this point clear throughout the current version of the manuscript.

(2) In the abstract you report that ROCK inhibition impairs the actin cytoskeleton around the skeleton. In examining your images in Fig. 5 that is not the case. Based on Phalloidin staining, actin surrounds both the control and the ROCK-inhibited skeleton. The distribution of actin is the same in both cases. Myosin is also stained in this figure and it too shows similar staining both in experimental and control. So, to this reviewer, there is insufficient evidence to suggest that the actin cytoskeleton is impaired, and there is no evidence directly relating ROCK with that cytoskeleton. I'm not questioning the observation that inhibition of ROCK causes stunting and mispatterning of the skeleton. That you show and quantify well. The issue is the precise target of ROCK. Your data does not establish the specific cause. It could be the actin cytoskeleton but your experiments do not directly address that.

Fig. 5 shows a clear difference between F-actin in control and under ROCK inhibition. In control F-actin is enriched around the spicule and under ROCK inhibition the spicule doesn’t form and disorganized F-actin is accumulated in the skeletogenic cells. Yet, as we stated above – this is not a proof for the direct effect of ROCK on F-actin polymerization, and we explain it explicitly in the results, lines 324-326 and in the discussion, lines 405-408.

(3) In parts of the manuscript you use the term filopodia and in other parts I think you use pseudopodia to refer to the same structure. Since Ettensohn has provided the most evidence on the organization of the skeletogenic syncytia, I suggest you use the same term he used for those cellular extensions.

The filopodia and the pseudopodia are two distinct structures generated by the skeletogenic cells. The filopodia is the common cellular extension described in many cells, while the term “pseudopodia cable” describes the specific structure that forms between the skeletogenic cells in which the spicule cavity forms, in agreement with Prof. Ettensohn terminology.

(4) In trying to find relationships you cite a number of previous papers at the end of the introduction. I went back to those papers and they describe (from your work) calcium exocytosis, plus filopodia formation, plus planar cell polarity, plus CDC42, any one of which could involve an actin cytoskeleton. You even cite a paper saying that perturbations of ROCK prevent spicule formation. I went back to that paper and that isn't the case. You then summarize the Introduction by relating ROCK and the actin cytoskeleton, thereby raising reader expectation that the two will be connected. As above, in reality, your evidence here does not connect the two.

We thank the reviewer for giving us credit for all these works, but only the paper on vesicle kinetics is from our lab (winter et al 2021). As for Croce et al, 2006 that the reviewer refers to: in Fig. 9A, 75µM of Y-27632 is used to inhibit ROCK in the same sea urchin species that we use, and the phenotype is identical to what we observe – the skeletogenic cells are there, but the spicule is not formed. As mentioned above, in the current version we distinguished clearly between our solid findings and our interpretations.

(5) You emphasize in Fig. 1 the inhibition of ROCK in the presence of VEGFR inhibition. However, at no place in the manuscript do you say anything about how VEGFR is inhibited, when it is inhibited, or how you know it is inhibited. That oversight must be corrected. You mention axitinib but don't say anything about what it does. Some readers may know its activity but many will not.

We now indicate that we use Axitinib to block VEGFR in the results section (line 104) and in the methods section (lines 470-471).

(6) Fig. 2. The use of Y27632 as a selective inhibitor of ROCK. According to data sheets from the manufacturer, at the levels used in your experiments, 120 µm, 80 µm and 30 µm, those levels of inhibitor also inhibit the activity of PKA and PKC (both inhibited at around 25 µm). This is concerning because of the literature indicating that activation of the VEGFR operates through PKA. Inhibition of PKA, then, would inhibit the activity of VEGF signaling. Thus, the inhibitory effects of Y27632 may actually not be attributed specifically to ROCK. Furthermore, the heading of this section states that ROCK activity controls initiation, growth, and morphology of the spicule. Yet, even in high levels of inhibitor spicule production is initiated. Yes, the growth and the morphology are compromised, but the initiation doesn't seem to be.

The spicule fails to form under ROCK continuous inhibition in all concentrations (Fig. 2). Also, as we explained in details above, these Ki values are based on biochemical experiments with purified proteins and are not relevant to in-vivo use of the inhibitor. Yet, these Ki values demonstrate that the affinity of the inhibitor to ROCK is 100 higher than of its affinity to PKA and PKC. Specifically to the reviewer suggestion here: direct inhibition of PKA does not have skeletogenic phenotypes, not in whole embryos (1) and not in skeletogenic cell culture (2). Since we see the same skeletogenic phenotypes at low Y-27362 concentration and the genetic and pharmacological pertrubations of ROCK reconcile, we believe that these phenotypes can be atributed directly to ROCK.

(7) The synchrotron study is very nice with two points that should be addressed. Again, a high concentration of Y27632 was used giving a caveat on ROCK specificity. And second, the blue and green calcein pulses are very nice but the recent paper by the Bradham group should be cited.

We added a reference to Bradham recent paper on two calcein pulses (10).

(8) Fig. 5 is where an attempt is made to associate ROCK inhibition to alterations in actomyosin. Again, a high concentration of the inhibitor is used casting doubt on whether it specifically inhibits ROCK. However, even if the inhibition is specific to ROCK the images do not provide convincing evidence that ROCK activity normally is directed toward actomyosin. This is crucial to the manuscript.

As stated above, we addressed the specificity in this version and we modified the text to emphasize the correlation and not cuasation: Fig. 5 shows a clear difference between F-actin in control and under ROCK inhibition. In control F-actin is enriched around the spicule and under ROCK inhibition the spicule doesn’t form and disorganized F-actin is accumulated in the skeletogenic cells. Yet, as we stated above – this is not a proof for the direct effect of ROCK on F-actin polymerization, and we explain it explicitly in the results, lines 324-326 and in the discussion, lines 405-408.

(9) Again in Fig. 6 the inhibitor is used with the same concern about whether the effects noted are due to ROCK.

Fig. 6 is now Fig. 7 – the effect of ROCK on gene expression and as explained above, we addressed the specificity in this version.

(10) Lines 350-358. This interpretation falls apart without showing that the inhibitor is specific for ROCK as indicated above. Also, Fig. 5 is unconvincing in showing a difference in actin or myosin distribution in control vs ROCK inhibited embryos. Yes, the spicules are stunted, but whether actin or myosin have anything to do with that as a result of lack of ROCK activity is not demonstrated.

As stated above, we addressed the specificity in the revised version and we modified the text to emphasize the correlation and not cuasation: Fig. 5 shows a clear difference between F-actin in control and under ROCK inhibition. In control F-actin is enriched around the spicule and under ROCK inhibition the spicule doesn’t form and disorganized F-actin is accumulated in the skeletogenic cells. Yet, as we stated above – this is not a proof for the direct effect of ROCK on F-actin polymerization, and we explain it explicitly in the results, lines 324-326 and in the discussion, lines 405-408.

(11) Throughout, the manuscript spelling, grammar, and sentence structure will require extensive editing. The mistakes are numerous.

We did our best to correct the spelling and grammar. If we still missed some mistakes, we would be happy to further correct them.

References

(1) Mitsunaga K, Shinohara S, Yasumasu I. Probable Contribution of Protein Phosphorylation by Protein Kinase C to Spicule Formation in Sea Urchin Embryos: (sea urchin/protein kinase C/spicule formation/H-7/HA1004). Dev Growth Differ. 1990;32(3):335-42.

(2) Mitsunaga K, Shinohara S, Yasumasu I. Does Protein Phosphorylation by Protein Kinase C Support Pseudopodial Cable Growth in Cultured MicromereDerived Cells of the Sea Urchin, Hemicentrotus pulcherrimus?: (sea urchin/protein kinase C/spicule formation/phorbol ester/H-7). Dev Growth Differ. 1990;32(6):647-55.

(3) Su Y, Huang H, Luo T, Zheng Y, Fan J, Ren H, et al. Cell-in-cell structure mediates in-cell killing suppressed by CD44. Cell Discov. 2022;8(1):35.

(4) Kagawa H, Javali A, Khoei HH, Sommer TM, Sestini G, Novatchkova M, et al. Human blastoids model blastocyst development and implantation. Nature. 2022;601(7894):600-5.

(5) Canellas-Socias A, Cortina C, Hernando-Momblona X, Palomo-Ponce S, Mulholland EJ, Turon G, et al. Metastatic recurrence in colorectal cancer arises from residual EMP1(+) cells. Nature. 2022;611(7936):603-13.

(6) Becker KN, Pettee KM, Sugrue A, Reinard KA, Schroeder JL, Eisenmann KM. The Cytoskeleton Effectors Rho-Kinase (ROCK) and Mammalian DiaphanousRelated (mDia) Formin Have Dynamic Roles in Tumor Microtube Formation in Invasive Glioblastoma Cells. Cells. 2022;11(9).

(7) Segal D, Zaritsky A, Schejter ED, Shilo BZ. Feedback inhibition of actin on Rho mediates content release from large secretory vesicles. J Cell Biol. 2018;217(5):1815-26.

(8) Fischer RS, Gardel M, Ma X, Adelstein RS, Waterman CM. Local cortical tension by myosin II guides 3D endothelial cell branching. Curr Biol. 2009;19(3):2605.

(9) Narumiya S, Ishizaki T, Uehata M. Use and properties of ROCK-specific inhibitor Y-27632. Methods Enzymol. 2000;325:273-84.

(10) Descoteaux AE, Zuch DT, Bradham CA. Polychrome labeling reveals skeletal triradiate and elongation dynamics and abnormalities in patterning cue-perturbed embryos. Dev Biol. 2023;498:1-13.

Version 2

Abstract

Abstract

Biomineralization had apparently evolved independently in different phyla, using distinct minerals, organic scaffolds and gene regulatory networks (GRNs). However, diverse eukaryotes from unicellular organisms, through echinoderms to vertebrates, use the actomyosin network during biomineralization. Specifically, the actomyosin remodeling protein, Rho-associated coiled-coil kinase (ROCK) regulates cell differentiation and gene expression in vertebrates’ biomineralizing cells, yet, little is known on ROCK’s role in invertebrates’ biomineralization. Here we reveal that ROCK controls the formation, growth and morphology of the calcite spicules in the sea urchin larva. ROCK expression is elevated in the sea urchin skeletogenic cells downstream of the Vascular Endothelial Growth Factor (VEGF) signaling. ROCK inhibition leads to skeletal loss and disrupts skeletogenic gene expression. ROCK inhibition after spicule formation reduces spicule elongation rate and induces ectopic spicule branching. Similar skeletogenic phenotypes are observed when ROCK is inhibited in a skeletogenic cell culture, indicating that these phenotypes are due to ROCK activity specifically in the skeletogenic cells. Reduced skeletal growth and enhanced branching are also observed under direct perturbations of the actomyosin network. We propose that ROCK and the actomyosin machinery were employed independently, downstream of distinct GRNs, to regulate biomineral growth and morphology in Eukaryotes.

Introduction

Biomineralization is the process in which organisms from the five kingdoms of life, use minerals to produce shells, skeletons and teeth that protect and support them (1–4). Recent studies suggest that biomineralization evolved by the phylum specific co-option of ancestral gene regulatory networks (GRNs) that drove the construction of distinct organic scaffolds (4–9) and by the evolution of specialized sets of biomineralization proteins (3, 7, 10–13). This explanation is in line with the dissimilar GRNs and biomineralization proteins that drive biomineralization in different phyla (5, 12, 14). Yet, there are common cellular processes required for mineralization in highly diverse phyla (4), suggesting that distinct upstream GRNs might have recruited common cellular and molecular mechanisms to drive biomineralization.

The actomyosin network was shown to play a role in biomineralization in various Eukaryote models, from unicellular organisms to vertebrates’ bones and teeth (15–22). There is a tight association between actin filaments and the biomineralization compartment in diatoms (19) and a highly dynamic actin organization that forms the biomineralization compartment in foraminifera (17). Perturbations of actin polymerization result with severely deformed shells and inhibition of shell shedding in coccolithophores (15, 16) and diatoms (18). In vertebrates, Rho GTPases and Rho-associated coiled-coil kinase (ROCK), regulate chondrocytes, osteoblasts and odontoblasts differentiation and affect genes expression in these biomineralizing cells (20–22). However, the roles of actomyosin remodeling in controlling mineral deposition and shaping biomineral morphology, are still unclear.

Sea urchin larval skeletogenesis provides an attractive model to study the role of the actomyosin network in biomineralization. Sea urchin larval skeletons are made of two frameworks of interconnected calcite rods, termed “spicules” that are generated by the skeletogenic cells (6, 23). To make the spicules, the skeletogenic cells form a ring with two lateral skeletogenic clusters and fuse through their filopodia forming a pseudopodia cable that links them into a syncytium (24, 25). The mineral is concentrated in the form of amorphous calcium carbonate (ACC) inside intracellular vesicles (26, 27). The vesicles are secreted into the biomineralization compartment generated in the skeletogenic cell clusters forming two triradiate spicules (Fig. 8A, (26, 28)). The tubular biomineralization compartment, also called the spicule cavity, elongates within the pseudopodia cable by localized mineral deposition at the tip of each rod (29).

The GRN that controls sea urchin skeletogenesis is known in great detail and is highly similar to the GRN that controls vertebrates’ vascularization, suggesting a common evolutionary origin of these two tubulogenesis programs (6, 23, 30). As the spicules elongate, the expression of key regulatory and biomineralization related genes becomes restricted to the skeletogenic cells proximal to the growing tips, possibly to promote mineral deposition at these sites (31–33). Localized gene expression is regulated by signaling cues such as the Vascular Endothelial Growth Factor (VEGF) signaling (30–32, 34). However, how the skeletogenic GRN drives spicule formation and localized mineral deposition at the tips of the rods is poorly understood.

Previous works suggest that the actomyosin network is essential for sea urchin skeletal growth. Calcium bearing vesicles perform an active diffusion motion in the skeletogenic cells with a diffusion length that inversely correlates with the strength and activity of the actomyosin network (35). Actin filaments are formed around the spicule (35, 36) and F-actin signal is enriched at the tips of the growing skeletal rods in skeletogenic cell culture (37). Genetic and pharmacological perturbations of the GTPase, CDC42, prevent the formation of filopodia in the skeletogenic cells and inhibit spicule formation and elongation (38). Pharmacological perturbations of ROCK prevent spicule formation (39) and genetic perturbations of Rhogap24l/2 result with ectopic spicule splitting (6). Despite these characterizations, little is known on the role of the actomyosin machinery in regulating sea urchin biomineral growth and morphology.

Here we study the role of ROCK and the actomyosin network in the sea urchin Paracentraorus lividus (P. lividus). Our findings reveal the critical role of ROCK and the actomyosin network in multiple aspects of sea urchin biomineralization, suggesting a common use of these factors in Eukaryote biomineralization, downstream of distinct GRNs.

Results

ROCK is enriched in the skeletogenic cells depending on VEGF signaling

We sought to study the spatial expression of the ROCK protein and its regulation by VEGF signaling, a prominent regulator of sea urchin skeletogenesis (6, 30). The sequence of ROCK and especially, its functional domains are highly conserved between human and sea urchin (Fig. S1, (40, 41)). According to RNA-seq data measured in P. lividus, ROCK is a maternal gene that degrades after the maternal to zygotic transition and picks up again after hatching (>14 hours post fertilization (hpf), Fig. S2A). We used a commercial antibody that recognizes human ROCK protein to test ROCK expression in P. lividus embryos. We first used western blot to detect ROCK expression in control and under VEGFR inhibition at the time of spicule formation and during skeletal elongation (Axitinib treatment, ∼22hpf, 27hpf and 33hpf, Fig. S2D). The antibody detected a ∼150kDa protein (Fig. S2D) that is the predicted size for P. lividus ROCK protein (153kDa). VEGFR inhibition marginally increased the overall level of ROCK at 22hpf, but did not affect it at 27hpf and 33hpf (Fig. S2E). Yet, this measurement was done on proteins extracted from whole embryos, and we wanted to study the spatial expression of ROCK and specifically, its regulation in the skeletogenic cells.

We studied the spatial distribution of ROCK protein using ROCK antibody and the skeletal cell marker, 6a9, that marks the skeletogenic syncytium membrane, in control and under VEGFR inhibition ((42), Fig. 1). We quantified ROCK signal in the skeletogenic cells compared to the neighboring ectodermal cells in both conditions, at the three time points mentioned above. In the three time points, under VEGFR inhibition, the skeletogenic cells do not migrate to their proper positions in the anterlateral and post-oral chains, as previously reported (6, 30, 34). At 22hpf, ROCK expression in the skeletogenic cells is mildly enriched compared to the neighboring ectodermal cells and this enrichment is not observed under VEGFR inhibition (Fig. 1A-F). At 27hpf, ROCK enrichment in the skeletogenic cells increases, but similar enrichment is observed under VEGFR inhibition (Fig. 1G-L). At 33hpf, ROCK enrichment in the skeletogenic cells is most apparent (Fig. 1M-O) and depends on VEGF signaling (Fig. 1P-S). At both 27hpf and 33hpf in control embryos, ROCK seems localized near the skeletogenic cell membranes, which further supports its activation in these cells, since ROCK activation leads to its localization to the cell membranes in other systems (43). Overall, this data demonstrate that ROCK expression is elevated in the skeletogenic cells and this enrichment strengthens with skeletal elongation and depends on VEGF signaling.

ROCK enrichment in the skeletogenic cells increases with time and depends on VEGF signaling.

(A-R) ROCK immunostaining at different times points in control and under VEGFR inhibition (150 nM Axitinib). In each condition and time point, the left image shows ROCK immunostaining and the skeletogenic cell marker, 6a9 (A, D, G, J, M and P). The middle image shows ROCK immunostaining alone in the whole embryo (B, E, H, K, N and Q). The right image shows the enlargement of the white rectangle region of the middle image (C, F, I, L, O, R). Scale bar in whole embryo is 50µm and in enlargements is 10µm. E – ectoderm, SC – skeletogenic cells. (S) quantification of the ratio between ROCK signal/area in the skeletogenic cells compared to the ectodermal cells (see methods for details). Each box plot shows the average marked in x, the median, the first and the third quartiles (edges of boxes) and all experimental measurement (dots). Experiments were performed in three independent biological replicates and in each condition at least 33 embryos were measured. Statistical significance was measured using paired 2-tailed t-test where, * indicates P<0.05, ** indicates P<0.005, and *** indicates P<0.0005.

ROCK activity in the skeletogenic cells controls spicule initiation, growth and morphology

Next, we studied the role of ROCK in sea urchin skeletogenesis using genetic perturbations. We downregulated the expression of ROCK by the injection of two different translation morpholino anti-sense oligonucleotides (MASOs). ROCK MASO-1 matches the start of translation region, ROCK MASO-2 matches the 5’ region upstream the start of translation, and random MASO was used as a control. Embryos injected with either ROCK MASO-1 or MASO-2 show reduced ROCK signal at 33hpf, supporting the downregulation of the ROCK protein using these MASO’s (Fig. S3B, F, J, N, R, V). Embryos injected with ROCK MASO-1 or MASO-2 show strongly reduced skeletons (Fig. 2C, D) and embryos injected with ROCK MASO-2 show an additional phenotype of severely branched skeleton (Fig. 2B). The additional branching phenotype and the larger percentage of affected embryos in ROCK MASO-2 are probably due to its higher efficiency resulting from its more 5’ location relative to translation start codon (Fig. 2E). Thus, the genetic perturbations of ROCK expression indicate that sea urchin ROCK is important for skeletal elongation and normal branching pattern.

ROCK activity is essential for spicule formation, normal elongation and branching in whole embryos and in skeletogenic cultures.

(A-E) genetic perturbation of ROCK translation using two different MASOs observed at 2dpf. (A) control embryo injected with Random MASO. (B) embryo injected with ROCK MO-2 shows ectopic spicule branching. (C, D) embryos injected with ROCK MO-1 or MO-2 show spicule initiations. (E) summary of MASO injection phenotypes based on 4-6 independent biological replicates. (F-K) pharmacological perturbations of ROCK activity using 10µM and 80µM of the inhibitor Y27632 observed at 2dpf. (F) representative control embryo with normal skeletal rods, B, body; AL, anterolateral; PO, post-oral and MV, midventral. (G) complete skeletal loss in embryo treated continuously with 10µM ROCK inhibitor. (J) reduced skeletal growth and enhanced ectopic branching in embryo where 10µM ROCK inhibitor was added at 25hpf. (J) small spicules with enhanced ectopic branching in embryo where 80µM ROCK inhibitor was added at 25hpf. (K) summary of perturbation phenotypes based on 3-8 biological replicates for each treatment. See additional concentrations, phenotypes and summary in Fig. S4 and Table S1. (L-O) representative spicules from skeletogenic cell cultures in control and under 30µM Y27632 at 72hpf. (L) linear spicule in control culture, (M) Y27632 addition at 16hpf, before spicule initiation, completely blocks spiculogenesis. (N, O) Y27632 addition after spicule initiation, at 48hpf, reduces spicule elongation (N), and enhances branching (O). (P) quantification of spicule length in control and ROCK inhibition (>48hpf) at 72hpf. *P < 0.05, **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment, except from 120µm that was done in two biological repeats. Scale bars are 50µm. In L, N and O, the numbers at the bottom indicate the number of spicules that show this phenotypes (left) over all observed spicules (right).

To elucidate the dependence of ROCK phenotypes on ROCK activity level and identify its effect on different stages of skeletogenesis, we tested the skeletogenic phenotypes of different concentrations of ROCK inhibitor, Y27632, applied at different time points (Fig. S4A). Y27632 binds to the ATP site of ROCK’s kinase domain and prevents its activity (44). Y27632 affinity to ROCK is 100 times higher than its affinity to other kinases, such as PKA and PKC (45). The amino-acids to which Y27632 binds, are conserved in the sea urchin ROCK protein, supporting its specific inhibition of this protein (Fig. S1C). Y27632 had been used in cell cultures and live embryos from vertebrates to Drosophila, in concentrations between 10-100µM (45–47). In the sea urchin embryo a concertation of 75µM of Y27632 was reported to completely block skeleton formation (39) and a concentration of 100µM was shown to delay gut invagination (48). Therefore, we tested the range of 10-80µM Y27632 applied before or after spicule initiation (Fig. S4A).

Continuous ROCK inhibition beginning at egg fertilization, resulted with significant skeletogenic phenotypes that are dose dependent (Fig. 2, Fig. S4). Continuous ROCK inhibition using 80µM Y27632, did not affect skeletogenic cell migration but eliminated skeletal formation at 27hpf (Fig. S4B, C). At two days post fertilization (dpf), complete skeletal loss is detected under continuous ROCK inhibition in all concentrations, ranging from 15% of the embryos exposed to 10µM to 67% of the embryos exposed to 80µM Y27632 (Fig. 2G, K, in agreement with (39)). The rest of the embryos exposed to continuous ROCK inhibition show either spicule initiations or small spicules with enhanced ectopic branching, with percentage depending on Y27632 concertation (Fig. 2K, S4K). Except from the skeletogenic phenotypes, the overall embryonic development of embryos exposed to Y27632 in all concentrations seems normal (Fig. 2G, J). Similar results are observed when adding the inhibitor at 20hpf, with distribution of phenotypes that depends on the concentration (Fig. S4I, K). Importantly, skeletal loss, spicule initiation and ectopic branching were not observed under PKC or PKA inhibition, that resulted with much milder skeletogenic phenotypes (PKC) or no skeletogenic phenotype (PKA, (49, 50)), supporting the selective inhibition of ROCK by Y27632. Altogether, continuous ROCK inhibition results with severe skeletogenic phenotype ranging from complete skeletal loss to small spicule with ectopic branching, with ratios that depend on the inhibitor concentration.

The addition of the inhibitor at 25hpf, after spicule initiation, results in a majority of embryos showing ectopic spicule branching (Fig. 2H-K, S4F-H, K). The reduction of skeletal growth rate and ectopic branching can be observed a few hours after the addition of ROCK inhibitor (Fig. S5, 30µM). Washing the inhibitor at the highest concertation after 25hpf results in partial recovery of skeletogenesis, with normal or mildly delayed skeletons, indicating that ROCK inhibition is reversable (Fig. S4E, K).

Overall, the genetic and pharmacological perturbations of ROCK result with sever reduction of skeletal growth and enhanced skeletal branching, but only continuous ROCK inhibition leads to complete skeletal loss (Fig. 2G, K). Immunostaining of fertilized eggs clearly shows that ROCK protein is maternal, in agreement with our RNA-seq data (Fig. S2A-C). The injected MASO’s cannot interfere with the maternal ROCK protein whereas the inhibitor affects the activity of both maternal and zygotic ROCK, which could underlie the absence of skeletal loss in the genetic perturbations. Indeed, ROCK signal is only moderately reduced in ROCK MASO injected embryos at 33hpf (Fig. S3F, J, N), demonstrating the partial penetration of the MASO. Thus, our genetic and pharmacological perturbations reconcile, and indicate that ROCK activity is necessary for spicule formation, skeletal elongation and normal branching pattern.

To test if ROCK skeletogenic phenotypes are due to its activity specifically in the skeletogenic cells, we inhibited ROCK in a culture of isolated skeletogenic cells (Fig. 2L-P). The addition of ROCK inhibitor to the skeletogenic cell culture at 16hpf, before the spicules form, completely abolished spicule formation (Fig. 2M). The addition of the inhibitor at 48hpf, after spicule formation, resulted with significantly shorter spicules (Fig. 2L, N) and increased branching compared to control spicules (Fig. 2O). The reduction of the spicule length becomes more significant with increasing concentration of the ROCK inhibitor (Fig. 2P). Notably, skeletal loss and ectopic spicule branching were not observed in PKA or PKC inhibition in skeletogenic cell cultures (50), further supporting the selective inhibition of ROCK by Y27632. ROCK skeletogenic phenotypes in the isolated skeletogenic cell culture are similar to its phenotypes in whole embryos, verifying that ROCK activity in the skeletogenic cells is essential for spicule formation, normal elongation and prevention of branching.

ROCK inhibition reduces spicule volume, surface area and total length, but not thickness (SR-µCT)

To quantify the effect of ROCK inhibition on mineral deposition and skeletal shape we used synchrotron radiation micro-computed tomography (SR-µCT) that enables three-dimensional visualization and quantification of spicule morphology at 2 and 3dpf (Figs. 3, S6, 40µM Y27632, see methods for details). The aberrant morphology of ROCK inhibited spicules is visible at both developmental time points (Fig. 3A). The ectopic branching occurs both at the tips where tip-splitting is detected (arrowheads, Fig. 3A), as well as in the back, where small spicule extensions, or mineral ‘dripping’ are observed (arrow at 48hpf, Fig. 3A). This demonstrates that the regulation of mineral deposition is perturbed both at the tip and in the back of the spicules under ROCK inhibition.

SR-µCT studies of ROCK inhibited spicuels show reduction in skeletal volume, surface area and total length, but not thickness and 2-pulse calcein shows loss of tip-dominance.

(A) Exemplary 3D-renderings of control spicules (top, blue) and spicules where 40µM of ROCK inhibitor was added at 25hpf (bottom, orange), dissected at 48hpf and 72hpf. Arrowheads point to tip splitting and arrow at 48hpf point to spicule dripping at the back. (B) Spicule volume vs. area for control and ROCK-inhibited spicules at 48 and 72 hpf. Each data point represents a single spicule. (C-D) Frequency distributions for volume and surface area of control and ROCK-inhibited spicules at 48 hpf and 72 hpf (left and right violin plots, respectively). (E) Spicules’ total branch length and average thickness for control and ROCK-inhibited spicules at 48 and 72 hpf. (F-G) Frequency distributions for spicule lengths and thickness of control and ROCK-inhibited spicules dissected at 48hpf and 72hpf. (H, I) Calcein two-pulses experiment. Embryos were exposed to blue calcein at 29-30hpf, to green calcein at 44-45hpf and stained with FM4-64 membrane marker (red) prior to image acquisition at 48hpf. (H) Control embryo, (G) Embryo where 30µM of Y27632 was added at 25hpf. The experiments were done in three biological replicates and the numbers at the bottom indicate the number of embryos that show this phenotype out of all embryos scored.

When compared to control spicules, ROCK inhibition leads to a ∼2.5-fold reduction of the spicule volume and surface area in both 2 and 3dpf (Fig. 3B-D, see Table S2 for measured values and standard deviations and S3 for statistical analyses). The average rate of spicule growth between 48hpf and 72hpf in control embryos is 325.5µm3/hour, and reduces to 119.8µm3/hour under ROCK inhibition. Hence, the rate of mineral deposition is significantly reduced under ROCK inhibition.

To understand the specific effect of ROCK inhibition on spicule elongation vs. thickening, we compared the total length and the mean thickness of the spicules between control and ROCK inhibition (Fig. 3E-G). The total length of control spicules was on average about 2.5-times longer than the length under ROCK inhibition in both 2dpf and 3dpf (Fig. 3E, F, tables S2 and S3). In contrast to its effect on the spicule length, ROCK inhibition caused a minor increase of the spicule thickness at 2dpf and at 3dpf the thickness was not affected by the inhibition (Fig. 3E-G, Tables S2, S3). These SR-µCT measurements indicate that ROCK activity plays a crucial role in controlling the rate of mineral deposition at the tips of the rods and spicule elongation, and is not affecting spicule thickening.

ROCK activity does not control mineral uptake but is required for “tip-dominance”

To test the effect of ROCK inhibition on mineral intake, we used the fluorescent chromophore, calcein that binds to divalent ions including Ca+2 and is widely used to track the calcium pathway in biomineralization (6, 28, 33, 35). We measured the number of calcein stained vesicles per area as well as the number of pixels of calcein signal per area in the skeletogenic cells, in control and under ROCK inhibition at 1dpf and 2dpf (Fig. S7A-E). Continuous ROCK inhibition does not change the number of calcium bearing vesicles at 1dpf, but the number of calcein stained pixels significantly increases in this condition, possibly indicting that the vesicle volume is larger (Fig. S7F, G). At 2dpf, the number of calcein stained pixels is similar in control and in continuous ROCK inhibition, suggesting that the overall calcium content does not change in this condition (Fig. S7H, I). The number of calcein stained vesicles is however decreased under ROCK inhibition, possibly indicating that there are fewer vesicles with larger volume. Addition of ROCK inhibitor at 25hpf affects neither the number of calcein stained vesicles nor the number of calcein stained pixels in the skeletogenic cells. Together, these measurements show that ROCK is not required for the uptake of calcium into the cells. Therefore, the skeletal phenotypes occurring under ROCK inhibition are not related to a decrease in calcium uptake or shortage in intracellular calcium, but are due to defects in mineral processing within the cells and in mineral deposition.

To monitor the effect of ROCK inhibition on mineral deposition rate and distribution, we applied two pulses of different calcein dyes (51): blue calcein was added at 29hpf and washed after one hour, followed by green calcein added at 44hpf and washed after one hour. Hence, blue calcein marks the initial triradiate region and green calcein marks the edges of the spicules that were recently deposited (blue and green arrows in Fig. 3H, I). The dye FM4-64 was used to mark the cell membranes. Under ROCK inhibition, the green labeled region is much closer to the blue labeled region compared to control, due to the reduction in skeletal elongation rate, in agreement with our SR-µCT measurements. However, while in control embryos each body rod has a single tip stained in green calcein, under ROCK inhibition one of the body rods has two tips and both are stained in green. This indicates that mineral is being deposited in both edges, and the mechanism that prevents the growth of multiple tips in each rod, enabling “tip-dominance”, (in analogy to plant stem apical dominance), is disrupted under ROCK inhibition.

The activity of the actomyosin network is essential for normal spicule elongation and inhibition of ectopic branching

In other systems, major roles of ROCK are to control F-actin polymerization and MyoII activation (52–56), hence we wanted to directly test the role of these actomyosin components in sea urchin skeletogenesis. To directly inhibit F-actin polymerization we used Latrunculin-A (Lat-A), that prevents the polymerization of actin monomers (57) and was shown to be effective in sea urchin embryos (58). To directly inhibit actomyosin contractility we used Blebbistatin (Blebb), that interferes with MyoII ATPase activity, prevents actin-myosin interactions (59) and was shown to be effective in sea urchin embryos (60). To prevent interference with the early cell divisions where the actomyosin network plays a critical role (61–63), we added the inhibitors, individually or together, before or after spicule formation (20hpf and 25hpf, see methods for details). We tested the resulting phenotypes at 2dpf (Fig. 4A-G).

Actin polymerization and myosin activation affect skeletal growth and branching.

(A-F) representative embryos showing the effect of actomyosin perturbations at 2df. (A) control embryo (B) embryo treated with 2μM Blebbistatin >25hpf, (C) embryo treated with 2μM Blebbistatin >20hpf, (D) embryo treated with 2nM Latrunculin-A and 1.5μM Blebbistatin >25hpf, (E) embryo treated with 2nM Latrunculin-A >20hpf, (F) embryo treated with 2nM Latrunculin-A >25hpf, arrow pointing to the additional spicule rod. (G) Statistics of Latrunculin-A and Blebbistatin phenotypes, color code of phenotype is indicated in the representative images. Error bars indicate standard deviation. All treatments were conducted in at least three biological replicates and exact number of replicates and scored embryos are provided in table S4. (H-J) representative spicules recorded at 72hpf from (H) control skeletogenic cell culture, (I) skeletogenic cell culture were 2nM Latrunculin-A was added at 48hpf and (J) skeletogenic cell culture were 2μM Blebbistatin was added at 48hpf. (K) Quantification of spicule length in the different treatments at 72hpf **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment. Scale bars are 50μm.

Our results indicate that F-actin polymerization and to a lesser extent, actomyosin contractility, are essential for normal skeletal growth and inhibition of ectopic branching. The majority of the embryos where MyoII activation was inhibited show ectopic branching mostly at the tips of the rods, and a minor delay in skeletal growth (Fig. 4B, C, G). These phenotypes are quite similar to those observe in the addition of 10µM of ROCK inhibitor at 25hpf (Fig. 2H, K). The majority of the embryos where F-actin formation was inhibited show significant reduction of skeletal growth and severe ectopic skeletal branching (Fig. 4E, F and similar phenotypes to Fig. 4D, statistics in Fig. 4G). These phenotypes are quite similar to those observed in 10µM continuous inhibition of ROCK and the addition of 30µM of ROCK inhibitor at 20hpf (Fig. 2H, K). In some embryos, additional spicules form (Fig. 4F, G). The co-inhibition of F-actin polymerization and MyoII activity results in skeletogenic phenotypes that resemble, but are slightly stronger, than those of the inhibition of F-actin alone (Fig. 4D, G). Thus, the phenotypes resulting from the inhibition of MyoII activation resemble those of the late inhibition of ROCK at low inhibitor concertation. The phenotypes resulting from inhibition of actin polymerization resemble those of the continuous inhibition of ROCK in low inhibitor concertation and the late inhibition of ROCK at higher inhibitor concentrations, yet skeletal loss is only observed under ROCK continuous inhibition.

To test whether actin polymerization and myosin activity specifically in the skeletogenic cells underlies the observed skeletogenic phenotypes, we inhibit both in an isolated skeletogenic cell culture (Fig. 4H-K). We added the inhibitors after the spicule formed (>48hpf) and observed the phenotypes at 3dpf. Both inhibitors reduced skeletal growth, with Lantruculin-A having a significantly stronger effect than Blebbistatin (Fig. 4K), in agreement with the stronger skeletogenic phenotypes of Lantruculin-A in whole embryos. However, differently than in whole embryos, skeletal branching in both Lantruculin-A and Blebbistatin is not affected and is similar to the observed branching in control cultures. The absence of the branching phenotype in the skeletogenic cell culture could be due to the increased rigidity of the substrate that could compensate for the reduce actomyosin activity. Another option is that the branching phenotype in whole embryos is due to the reduction of the activity of the actomyosin network in non-skeletogenic cells. Nevertheless, these results indicate that normal skeletal growth depends primarily on F-actin polymerization and to a lesser extent, on myosin contractility in the skeletogenic cells.

ROCK activity is required for F-actin organization around the forming spicule

The similarity between the skeletogenic phenotypes under ROCK inhibition and the direct perturbations of the actomyosin network led us to test the effect of ROCK inhibition on F-actin organization and MyoII activity. To accomplish this we used phalloidin, a fluorophore that binds F-actin specifically, an antibody that detects phosphorylated MyoII (MyoIIP, (35)) and the skeletogenic marker, 6a9 (Fig. 5). In control embryos, F-actin is detected around the tri-radiate calcite spicules at 27hpf (green arrow in Fig. 5B, in agreement with (35, 36)). The pseudopodia cable that connects the skeletogenic cells shows lower F-actin signal compared to the signal around the spicule (blue line in Fig. 5B, C, E blue arrows in Fig. 5G, H, J). MyoIIP signal is not enriched in the skeletogenic cells nor in the spicule (Fig. 5D, I). Under continuous ROCK inhibition, the pseudopodia that connects the skeletogenic cells still forms (Fig. 5R) and enhanced F-actin signal is still observed in the lateral skeletogenic cell clusters (green arrow in Fig. 5L); but unlike control embryos, the spicule cavity is not formed and F-actin is not organized around it (Fig. 5L). MyoIIP signal seems unchanged by ROCK inhibition at this time (Fig. 5N, S), in agreement with the weak and late phenotype of MyoII inhibition (Fig. 4). Thus, the pseudopodia cable forms but F-actin organization around the spicules does not occur under continuous ROCK inhibition, however it is not clear if the effect on F-actin organization is direct or due to the absence of spicule in this condition.

ROCK inhibition effect on F-actin organization and MyoII activity at 27hpf.

Representative images showing normal embryo (A-J) and embryos treated with ROCK inhibitor from fertilization (K-T, 80µM), at 27hpf. Phalloidin (green) was used to stain F-actin, MyollP (red) was used to stain phosphorylated myosinII and 6a9 (blue) was used to mark the skeletogenic cells. Right panels show enlargements of the rectangle sections marked in the left most panels. Green arrows point to regions that show enriched F-actin signal, blue arrows point to the pseudopodia cable. The blue line in B, C and E marks a region of the pseudopodia cable that is stained by 6a9 but has low F-actin signal. The experiments were done in 3 biological replicates, the numbers at the bottom left of (A, K) indicate the number of embryos that show this phenotype out of all embryos scored. Scale bar in A, F, K and P is 50μm, and in B, I, N and S is 10μm.

F-actin is enriched at the tips of the elongating spicules independently of ROCK activity

To assess the role of ROCK in actomyosin organization during skeletal elongation we compared F-actin and myoIIP signals between control embryos and embryos where ROCK inhibitor was added at 25hpf (Fig. 6). At the prism stage, we did not detect a clear difference in MyoIIP signal in the skeletogenic cells, between control and ROCK inhibited embryos (Fig. 6D, I). At this time, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Figs. 6B, G) as well as in ROCK morphants (green arrowheads in Fig. S3K, W). In both control and ROCK inhibited embryos, the F-actin signal was markedly higher in the regions of the pseudopodia cable where the spicule cavity had formed, compared to the regions where the spicule cavity had not yet formed (blue arrows in Fig. 6B-E, G-J). We quantified the F-actin signal at the tips compared to the back in control and ROCK inhibited embryos (Fig. 6K, L, see methods for details). The phalloidin signal is on average 3-fold stronger at the tips of the spicules compared to the back, in both control and ROCK inhibition, indicating that F-actin enrichment at the tips is significant and independent of ROCK activity.

F-actin filaments are enriched at the tips of the spicules and the tip to back ratio is unaffected by ROCK inhibition

(A-J) representative images at 33hpf, showing normal embryo (A-E) and embryos treated with 30µM ROCK inhibitor from the gastrula stage (>25hpf, F-J). Embryos are stained with Phalloidin (green), MyollP antibody (red) and 6a9 (blue). (B-E, G-J) enlargement sections of the spicule area marked with rectangle in A and F. White arrowheads point to the enriched F-actin signal at the tips. Blue arrows point to the region of the pseudopodia cable that is not filled with the spicule cavity. (K, L) quantification of the tip to back F-actin signal (number of green pixels per area) at 33hpf in control embryos and ROCK inhibition >25hpf. Each box plot shows the average (x), median (middle line), the first and the third quartiles and all the points measured. Asterisks indicate statistical significance in paired t-test, where * is p<0.05 and *** is p< 0.0005, NS- not significant. (M-V) similar experiments to (A-J), at 48hpf, 40µM of ROCK inhibitor added at 25hpf. Red arrows point to non-skeletogenic cells enriched with MyoIIP. The experiments were repeated in 3 biological replicates and the numbers at the bottom left of (A, F, M, R) indicate the number of embryos that show this phenotype out of all embryos scored. Scale bar in A, F, M, R is 50µm and in B, G, N, S, is 10µm. (W, X) representative spicules out of three biological replicates from control skeletogenic cell culture (W, n=60), and skeletogenic cell treated with 30µM ROCK inhibitor added at 48hpf and recorded at 72hpf (X, n=52). Left panel is phase image, middle panel is phalloidin staining and right panel shows the overlay. Green arrows point to the enhanced F-actin signal at the tips. Scale bar is 20μm.

At the pluteus stage, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Fig. 6N-Q, S-V). Some non-skeletogenic cells that are enriched with MyoIIP signal are detected at this time (red arrowheads in Fig. 6N-Q). Together these data demonstrate that F-actin coats the spicule cavity and is enriched at the tip of the rods, independently of ROCK activity.

We tested the effect of ROCK inhibition on F-actin in skeletogenic cell cultures. F-actin is detected at the tips, in both control cultures and cultures where ROCK was inhibited after spicule formation (Fig. 6W, X). Yet, under ROCK inhibition, branching is enhanced and F-actin is enriched at the splitting tips of the spicule rods (Fig, 6X). The enrichment of F-actin at the splitting tips of the spicule under ROCK inhibition in skeletogenic cell culture resembles the calcein staining at the two tips in the embryo in this condition (Fig. 3I). These observations further support the role of ROCK activity in regulating tip-dominance during sea urchin skeletogenesis.

ROCK activity is essential for normal skeletogenic gene expression

The role of ROCK in regulating gene expression during vertebrates’ chondrocytes, osteoblasts and odontoblasts differentiation (20–22) intrigued us to study the effect of ROCK inhibition on skeletogenic gene expression in the sea urchin embryo. ROCK continuous inhibition significantly downregulates the expression of multiple skeletogenic genes, including the cytoskeleton remodeling gene, rhogap24l/2, the biomineralization genes, caral7, scl26a5 and SM30 at 27hpf and 2dpf (Fig. 7A, (6)). ROCK inhibition did not affect the expression level of VEGFR and ROCK itself at these times (Fig. 7A). Washing the inhibitor after 25hpf or adding the inhibitor at 25hpf had a weaker effect, but still led to downregulation of caral7, SM30, angio1 and notch1 (Fig. 7B). Thus, ROCK activity is required for the normal expression level of multiple skeletogenic genes during sea urchin skeletogenesis.

ROCK activity is essential for normal gene expression in the skeletogenic cells.

(A, B) the effect of 80µM ROCK inhibition on gene expression. (A) continuous ROCK inhibition at 27hpf and 50 hpf (n=4). (B) addition of ROCK inhibitor at 25hpf and the wash of ROCK inhibitor at 25hpf, measured at 50hpf (n=3). Asterisks indicate p<0.001, one tailed z-test. Error bars show standard deviation. (C-F) Representative images of control embryo (top panels), embryos where ROCK inhibitor was added at 25hpf (middle panels), and embryos that were exposed to continuous ROCK inhibition (bottom panels), at the pluteus stage (∼48hpf). (C) skeletogenic cell marker, 6a9. MV, midventral; AL, anterolateral, and PO, Post-oral rods, D, dorsal skeletogenic chain. (D-F) spatial expression of skeletogenic genes. Gene names are indicated at the top of each panel. Numbers at the bottom of each image indicate the number of embryos that show this phenotype (left) out of all embryos scored (right), conducted in at least three independent biological replicates.

Both continuous and late ROCK inhibition strongly affect the spatial expression of key regulatory and biomineralization genes that are normally expressed differentially within the skeletogenic lineage at the pluteus stage (Figs. 7D-F, S8). These include, the VEGF receptor (VEGFR), the biomineralization genes SM50 and SM30, skeletogenic transcription factors, Ets1, Alx1 and Erg1 that are essential for skeletogenic cell specification and biomineralization (23, 64) and the transcription factors Hex and MyoD1. In normal plutei these genes are differentially expressed within skeletogenic lineage, but in both continuous and late ROCK inhibition the expression of these genes becomes localized to the two lateral skeletogenic cell clusters and their vicinity (Fig. 7D-F, S8). The lateral skeletogenic cells clusters are the cells where these genes are expressed in the gastrula stage (24-27hpf) and where spicule formation begins in normal embryos (6, 32). The expression of VEGFR and SM50 under ROCK inhibition is more expanded than the other tested genes and is also observed at the anterolateral rods (Fig. 7D, E). The VEGF ligand that is normally expressed in the ectodermal cells near the four tips of the post-oral and anterolateral rods (32), is expressed at four ectodermal domains near the lateral skeletogenic cell clusters and near the end of the anterolateral chain, under ROCK inhibition (Fig. S8G). Our data shows that under ROCK inhibition, the expression of key skeletogenic genes remains localized to the area near the lateral skeletogenic cells clusters (Fig. 7D-F, S8), despite the proper formation of the dorsal, anterolateral and mid-ventral chains (Fig. 7C). Overall, ROCK activity is essential for normal expression level and spatial gene expression in the skeletogenic lineage.

Discussion

Biomineralization is a complex morphogenetic process, encoded in the species genome and executed by the GRNs and the cellular machinery. Here we used the sea urchin larval skeletogenesis to investigate the role of ROCK and the actomyosin network in biomineral formation and in regulating gene expression, as we summarize in Fig. 8. ROCK expression is enriched in the skeletogenic cells, partially depending on VEGF signaling (Fig. 1, 8C). ROCK activity in the skeletogenic cells is necessary for spicule formation, skeletal elongation and prevention of ectopic skeletal branching (Figs. 2-3, S4, S5). Direct inhibition of F-actin polymerization results in similar but not identical skeletogenic phenotypes (Fig. 4). F-actin is organized around the spicule cavity and enriched at the spicule tips (Figs. 5, 6). ROCK activity feeds back into the skeletogenic GRN and affects regulatory and biomineralization gene expression (Fig. 7). Below we discuss our findings and their implications on the biological control and evolution of biomineralization.

The role of ROCK and the actomyosin network in sea urchin skeletogenesis, summary.

(A) the spicule forms at about 24hpf in P. lividus and at 27hpf, the triradiate spicule is coated with F-actin. (B) ROCK activity is required for spicule initiation and for F-actin organization around the spicule. (C) model of the functional links between the skeletogenic GRN, ROCK, F-actin and their skeletogenic outcomes, during spicule initiation (∼24-27hpf in P. lividus). Cian arrows indicate links discovered in this work. (D) during skeletal elongation F-actin is detected around the spicule cavity and is enriched at the tips of the rods. The expression of SM50 and some other skeletogenic genes, is localized to the tips of the rods. (E) under the addition of ROCK inhibitor, skeletal growth rate is reduced and ectopic spicule branching is observed. The expression of skeletogenic genes is localized to the vicinity of the growing rods. (F) the functional links between the skeletogenic GRN, ROCK, F-actin and MyoII and their skeletogenic outcomes during skeletal elongation. Cian arrows indicate links discovered in this work.

The first step in generating the sea urchin spicules is the construction of the spicule cavity, a membrane filled with calcium carbonate and coated with F-actin (Fig. 8A). ROCK activity is necessary for the spicule initiation and for the organization of F-actin around the spicule cavity (Figs. 2, 5, 8B). However, since the spicule doesn’t form under ROCK inhibition, it is hard to conclude if the absence of F-actin organization around the spicule is due to ROCK regulation of actin polymerization or due to the absence of spicule in this condition. Relatedly, a direct perturbation of F-actin polymerization strongly disrupts spicule growth and shape, but not spicule formation, indicating that F-actin is not essential for spicule initiation (Fig. 4). We therefore propose that ROCK activity is necessary for spicule initiation, but not through its’ direct regulation of the actomyosin network, but though its other targets, yet to be identified (Fig. 8C).

After the spicule cavity has formed, the spicules elongate inside the pseudopodia chord by mineral deposition at the tips of the spicule rods, which depends on ROCK activity, on F-actin polymerization and to a lesser extent, on MyoII contractility (Figs. 2-4, 8D-F). F-actin is detected along the spicules and is significantly enriched at the tips of the elongating spicules, independently of ROCK activity (Fig. 5, 6, 8D, E (65)). Under ROCK inhibition and under the inhibition of F-actin polymerization, ectopic branching is observed at the tips and in the back (Figs. 2H, J, O, 3A, I, 4, 8E). MyoII inhibition results with ectopic branching at the tips of the rods (Fig. 4B, G), that is similar to the effect of late ROCK inhibition at low concertation (Fig. 2H, K). These correlative similarities between ROCK and the actomyosin perturbations lead us to the following speculations: the low dosage of late ROCK inhibition is perturbing mostly ROCK activation of MyoII contractility while the higher dosage affects factors that control actin polymerization (Fig. 8F). Further studies in higher temporal and spatial resolution of MyoIIP activity and F-actin structures in control and under ROCK inhibition will enable us to test this.

The participation of ROCK, F-actin and MyoII in polarized growth and vesicle exocytosis has been observed in both animals and plants (65, 66). F-actin is used for short range myosin trafficking of vesicles toward the secretion site (65, 67). Once the secretory vesicle fuses with the membrane, it is coated with F-actin that stabilizes the vesicle and enables MyoII binding and ROCK-regulated contraction, necessary for content secretion (46, 47). In plant cells, F-actin accumulate at the growing tip of the cell and assists vesicle exocytosis necessary for polarized tip-growth (66). ROCK, F-actin and MyoII could be regulating the localized exocytosis of mineral-bearing vesicles during sea urchin skeletogenesis. The reduction of spicule growth rate and the enhanced branching could be due to impaired and misplaced vesicle trafficking and deposition under ROCK and F-actin inhibition. Detailed investigation of the kinetics and deposition of mineral-bearing vesicles and the role of actomyosin network in these processes will illuminate the regulation of vesicle exocytosis during biomineralization.

ROCK activity is necessary for correct spatiotemporal expression of regulatory and biomineralization genes (Fig. 7). ROCK inhibition downregulates the expression level of skeletogenic genes (Fig. 7A) and prevents the spatial progression of key skeletogenic genes toward the edges of the skeletal rods (Fig. 7, S8, 8E). This implies that ROCK activity provides regulatory cues to the skeletogenic GRN required for the dynamic progression of gene expression within the skeletogenic lineage. In other systems, ROCK activity was shown to regulate gene expression through various intracellular pathways (68–70). Furthermore, ROCK has an important role in mechano-sensing of matrix rigidity and transducing it into changes in gene expression (71). Relatedly, stiff substrate activates ROCK in pre-osteoblastic cells in-vitro, which activates Erk that leads to osteogenic differentiation and upregulation of osteogenic gene expression (70). Sea urchin ROCK could be a part of the mechano-sensing mechanism that detects the high spicule stiffness and transduces it into gene expression, providing a necessary feedback between spicule elongation and the skeletogenic GRN.

Overall, our findings together with the role of the actomyosin network in biomineralization, from calcifying and silicifying single cells to vertebrates’ bones and teeth (15–22, 70), suggest that this molecular machinery is a part of the common molecular tool-kit of biomineralization. Most likely, the actomyosin network was employed independently, downstream of distinct GRNs across Eukaryotes, to control mineral growth and morphogenesis.

Methods

Animal and embryos

Adult Paracentrotous lividus were obtained from the Institute of Oceanographic and Limnological Research (IOLR) in Eilat, Israel. Eggs and sperm were obtained by injection of 0.5M KCl solution into the adult sea urchin species. Embryos were cultured at 18°C in 0.2µ filtered ASW.

Imaging

Embryonic phenotypes and WMISH were imaged by Zeiss Axioimager 2. Fluorescent markers were imaged using Nikon A1-R Confocal microscope. All images were aligned in Adobe photoshop and Illustrator.

VEGFR inhibitor, Axitinib (AG013736) treatment

Axitinib (AG013736, Selleckchem, Houston, TX, USA) was applied as described in (32)

Western Blot

Embryo cultures treated with 150 nM Axitinib, or DMSO as control, were collected at 22hpf, 27hpf and 33hpf by centrifugation of 10,000g for 2 minutes. We lysed the pellet in 150µL lysis buffer (20 mM Tris-HCl, 150 mM NaCl, 1% Triton X-100, pH 8) with protease inhibitors (protease inhibitor cocktail w/o metal chelator; Sigma, P8340; 1:100) and phosphatase inhibitors (50 mM NaF, 1 mM Na3VO4, 1 mM Na4P2O7, 1mM, βGP) as described in (72). 40µg protein were loaded on 8% SDS-acrylamide gel, transferred to PVDF membranes for Western blot analysis. Anti-ROCK antibody (ab45171, Abcam) was used in a 1:300 dilution followed by anti-rabbit HRP secondary antibody diluted to 1:2500 (Jackson ImmunoResearch 111-035-003). For the loading control, the membrane was incubated with anti-β-tubulin antibody (1:5000; Sigma) followed by anti-mouse antibody 1:5000 (Jackson ImmunoResearch 115-035-003). Quantification of ROCK signal was done using Image studio lite vr. 5.2. ROCK signal was divided by Tubulin signal and then the ratio between ROCK/Tubulin in control vs. VEGFR inhibition was calculated. The graph in Fig. S2 was generated based on three biological replicates (different sets of parents) at 27hpf and 33hpdf and four biological replicates at 22hpf.

Immunostaining procedure

Phalloidin labeling, MyoIIP (p20-MyosinII), ROCK (anti ROCK2+ROCK1 antibody [EP786Y], AB-ab45171, Abcam), and 6a9 immunostaining were done similarly to (35).

Quantification of Anti-ROCK and Phalloidin signal

A graphical program was written to allow manual quantification of the fluorescent signal from the spicule regions by identifying “stained” pixels per selected area. The ratio of stained anti-ROCK to region total area was compared between ectodermal and skeletogenic cells, and between control and VEGFR inhibited embryos at 22, 27 and 33hpf. The average ratios and ztest for significance difference from 1, are 22hpf control, ∼1.3, z=0.0006; VEGFR inhibition ∼1, z=0.5, 27hpf control ∼1.8, z<10-11; VEGFR inhibition ∼1.6, z<10-8, 33hpf control ∼4.6, z<10-23; VEGFR inhibition ∼1.5, z<10-5. For phalloidin analyses, embryos at 33hpf of control and ROCK inhibitor addition at 25hpf were used and for each image, two regions were selected: an area at the tip of the spicule and an area at the back of the spicule, about ∼10 microns apart. The ratio of stained phalloidin to region total area was compared between tip and back, in control and ROCK inhibited embryos. Paired t-test was used to compare the differences in phalloidin signal between the tip and back of the spicules across all groups and between control and ROCK inhibited embryos.

ROCK MASOs injections

Microinjections were conducted similarly to (32). The eggs were injected with an injection solution containing 0.12M KCl, 0.5µg/µl Rhodamine Dextran and 900µM of MASOs. Translation MASOs were designed and synthesized by Gene Tools, Philomath, OR, according to ROCK sequence (40). The sequence of ROCK MASO-1 (5’- AGACATATTTGGAGCCGA[CAT]CCTG-3’) matches the start of translation region and ROCK MASO-2 (5’-TCTCTTGCGTTATATTCCACTAAGT-3’) matches the 5’ region upstream of the start of translation (40). Control embryos were injected in the same concentration with Random MASO. Injected embryos were cultured at 18°C, imaged and scored at 2dpf.

ROCK inhibitor (Y27632) treatment

Y27632 stock (10005583 CAS Regisrty No. 129830-38-2, Cayman chemical), was prepared by reconstituting the chemical in DMSO. The embryos were treated with Y27632 at a final concentration between 30-80µM, as mentioned in the results. Throughout the paper, control embryos were cultured in DMSO at the same volume as Y27632 solution and no more than 0.1% (v/v).

ROCK inhibition in isolated skeletogenic cell culture

Skeletogenic cell culture was performed as described in (37) with minor changes. Isolated skeletogenic cells were cultured in CultureSlides (4 chambers polystyrene vessel tissue culture treated glass slide REF 354104) with ASW (Red Sea Fish LTD) containing Gentamicin and Penicillin-Streptomycin (GPS) at 18°C. Media containing 4% (v/v) horse serum (Sigma-Aldrich H1270) in ASW+GPS was added to the culture. ROCK inhibitor, Y27632, was added to the cell culture at 16hpf (before spicule initiation) or 48hpf (after spicule initiation), and images were taken at 72hpf. Each treatment was conducted in three biological repeats except from the >48hpf addition of 120µM which was done in two biological repeats.

Quantification of skeletal length and statistical analysis

Skeletal length measurement was done as described in (32). The measurements were repeated for three biological repeats for control, ROCK inhibition with 30μM and 80μM, Latrunculin-A 2nM, Blebbistatin 2µM, and 2 biological repeats for 120μM. In ROCK inhibition experiments, a total of 275 skeletons were measured for control, 105 skeletons for 30μM, 122 skeletons for 80μM and 80 for 120μM. In Latrunculin-A and Blebbistatin experiments, a total of 116 skeletons were measured for control, 149 for Latrunculin-A and 107 for Blebbistatin. The data were analyzed in Excel and the statistical analysis was performed using Kruskal-Wallis non-parametric test in Excel as described in (73) and SPSS statistics 27.

Spicule preparation for micro-computed tomography (SR-µCT) measurement

Control embryos and embryos where 40µM Y27632 was added at 25hpf, were collected by centrifugation of 50mL Falcon tubes at 2000rpm, at 48h and 72hpf. Every tube was washed in 10 mL cold distilled water (DW), transferred into epi-tube and washed three times in 1mL cold DW (incubation at 4°C between every wash, centrifuge at 2000rpm in room temperature for 2min). Then the skeletons were washed three times in 3% NaOCl followed by three washes in DW, one wash in 70% EtOH and one wash in Acetone. Skeletal spicules were dried at room temperature and stored at -20 °C.

Synchrotron and lab-based SR-µCT

Dry and loose skeletal spicules prepared as described above, were used for the SR-µCT analysis. We analyzed 14 sets of spicules from three (control) and four (ROCK inhibited) pairs of parents, divided into four groups: 2dpf, control, n=44; 2dpf ROCK inhibited, n=93; 3dpf control, n=51; 3dpf ROCK inhibited, n=93, see Table S1). Spicules were glued to sharpened toothpick tips, either in bulk or individually when picked by hand by means of a brush bristle. Each toothpick loaded with spicules, was mounted on a 3ml vial twist-off lid and for transportation securely stored in the vial (Fig. S5A, B). For each set 2-3 µCT-samples were prepared, which allowed the acquisition of statistically relevant sample sizes (40+ spicules/group, Table S1). Tomographic data sets were acquired using the following scanning parameters; for the lab based µCT (RX Solutions, Chavanod, France) 90kV, 143-160µA, 1440 images, average frames 10, frame rates 0.4-1, voxel sizes: 0.45-59µm; and for the synchrotron radiation µCT (Anatomix, Synchrotron SOLEIL, Gif-sur-Yvette, France) 2048 images, angle between projections 0.09°, voxel size: 0.65µm. Tomographic data reconstruction was performed with commercial (RX Solutions, Chavanod, France) and PyHST2 software (A. Mirone et al., Nucl. Instrum. Meth. B 324 (2014) 41-48, doi:10.1016/j.nimb.2013.09.030) at the MPICI and at Anatomix, SOLEIL, respectively.

Three-dimensional (3D) data analysis

Data visualization, pre-processing including cropping and image type conversion to 16bit (of the Sr-µCT data), and user-augmented segmentation of intact spicules were performed in Amira using off-the-shelf modules (Fig. S5C-E, (74)). Volume and area measurements were performed on the segmented spicules, i.e. the label-fields using Amira’s material.statistics module (Volume3D and Area3D). Average spicule thickness and total length measurements were performed with python code, utilizing available python libraries (e.g., skimage, skan, scipy). The code was applied to 3D-tif files created from Amira’s label-fields, which contained the segmented spicules. For length measurements, the spicules were skeletonized and the chord lengths of the branches forming a spicule-skeleton were summed up to obtain the total spicule length. For thickness measurements, we performed Euclidean distance transforms for each spicule using scipy’s ndimage.morphology.distance_transform_edt package, and calculated the average spicule thickness based on the distance measurements obtained at each voxel from spicule-skeleton to the nearest spicule surface. Quality control of our measurements was performed with open source software, Fiji and available plugins (e.g. BoneJ, FiberJ, DiameterJ).

Calcein staining

A 2mg/ml stock solution of blue calcein (M1255, Sigma) and a 2mg/ml stock solution of green calcein (C0875, Sigma) were prepared by dissolving the chemicals in distilled water. Working solution of 250μg/ml and 150μg/ml, respectively, was prepared by diluting the stock solution in artificial sea water. Blue calcein was added to the embryo culture at 29hpf for one hour and then washed. At 44hpf the green calcein was added for one hour and washed. For FM4-64 staining A 100µg/ml stock solution of FM4-64 (T13320, Life technologies, OR, USA) was prepared by dissolving the chemical in distilled water. Working solution of 2.5 µg/ml was prepared by diluting the stock solution in artificial sea water. Embryos were immersed in working solution about 10 minutes before visualization.

Latrunculin-A and Blebbistatin treatments

1mM of Lat-A stock (L12370, Thermo Fisher Scientific) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Lat-A at final concentration of 2 nM. 1mM of Blebbistatin stock (13013-5, Enco) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Blebbistatin at final concentration of 2 µM. For the double inhibition experiments, embryos were treated with Blebb at final concentration of 1.5 µM, and with Lat-A at final concentration of 2 nM.

WMISH procedure

WMISH was performed as described in (32). List of the primers used to make the WMISH probes is available in (32).

cDNA preparation for QPCR experiments

For ROCK inhibition experiments total RNA was extracted from >1000 sea urchin embryos in each condition using RNeasy Mini Kit (50) from QIAGEN (#74104) according to the kits’ protocol. DNase treatment on column was done using RNase-Free DNase Set-Qiagen (50) (#79254). RNA was reverse-transcribed using the High Capacity cDNA RT kit, AB-4368814 (Applied Biosystems) following the manufacturer’s instructions.

Quantitative polymerase chain reaction (QPCR)

QPCR was carried out as described in (32). Complete list of primer sequences used for the experiments in provided there.

Supporting information

eLife Assessment

This valuable study addresses the role of Rho-associated coiled-coil kinase (ROCK) and the cytoskeleton in the initiation and growth of the calcified endoskeleton of sea urchin embryos. Perturbation by two independent approaches (a morpholino and a selective inhibitor) provide convincing evidence that ROCK participates both in actomyosin regulation and in the gene regulatory network that controls skeletogenesis. Exciting areas of future work will be to elucidate the mechanisms by which ROCK influences gene expression and to further dissect the role of the cytoskeleton in mineralization.

Reviewer #1 (Public Review):

In their revised manuscript Hijaze et al. adequately addressed the majority of my previous concerns in a satisfactory manner. In particular, they validated their morpholino knock-down experiments by explaining how they determined the optimal concentrations and provided an immunohistological evidence for the reduction in ROCK protein abundance. The authors also added new antibody stainings providing evidence that ROCK and F-actin do not interact directly but likely through other kinases that modulate f-actin, and that the localization of f-actin at the spicule tips remains unaffected by the knock-down. In addition, the authors revised their discussion to not overstate their observations, and by focusing on the potential mechanisms by which ROCK may affect biomineralization (i.e. mechano sensing and exocytosis of vesicles). Here I would like to add, that f-actin mediated exocytosis does not necessarily target mineral baring vesicles but may also promote the exocytosis of matrix proteins that are essential for the normal formation of the spicules and that are an integral component of other biominerals, as well. I strongly encourage the authors to continue on this exciting research, including the development of methods to analyze the molecular mechanisms that control vesicular trafficking in mineralizing systems.

Reviewer #2 (Public Review):

This project is on the role of ROCK in skeletogenesis during sea urchin development. That skeleton is produced by a small number of cells in the embryo with signaling inputs from the ectoderm providing patterning cues. The skeleton is built from secretion of CaCO3 by the skeletogenic cells. The authors conclude that ROCK is involved in the regulation of skeletogenesis with a role both in regulating actomyosin in the process, and in the gene regulatory network (GRN) underlying the entire sequence of events.

The strength of the paper is that they show in detail how perturbations of ROCK results in abnormal actomyosin activity in the skeletogenic cells, and they show alterations both in expression of transcription factors of the GRN, and expression of genes involved in assembly of the skeletal matrix. Two different approaches lead to this conclusion: morpholino perturbations and the actions of a selective inhibitor of the kinase activity. Thus, they achieved their goal which was to test the hypothesis that ROCK is involved in the process of skeletogenesis. Those tests support the hypothesis with data that was quantitatively significant.

The discussion was transparent regarding where the analysis ended and where the next phase of work should begin. While actomyosin involvement was altered when ROCK was perturbed, it isn't known how direct or indirect the role of ROCK might be. Also, while the regulatory input to spicule initiation and growth is affected when ROCK is inhibited, it isn't clear exactly where ROCK is involved.

Author Response

The following is the authors’ response to the previous reviews.

Thank you and the two reviewers for the thorough review of our manuscript. We thank you very much for the positive evaluation of our manuscript and your encouragement to continue in this fascinating topic. In this version we made minor changes in the text to address the comments and suggestion of the second reviewer and increase the clarity of the text.

Reviewer #2 Recommendation to the authors

We thank the reviewer for the sharp comments that help us improve the clarity of the paper. Below we list the changes we made to correct and revise the paper in accordance to the reviewer’s comments.

(1) Line 90. Isn't the genus Paracentrotus?

Yet it is, thank you. We corrected the typo.

(2) Figure 1 and supplementary figure 2. To this reviewer supplementary Figure 2 doesn't really help the story as written in the paragraph from line 96-110. You want to report expression of ROCK in skeletogenic cells. You do that quite well in Figure 1. Since Fig. S2 reports whole embryo expression of ROCK when only 5% of the cells in the embryo are the subject of interest here, and the Axitinib is selective, presumably for skeletogenic cells, the relative lack of effect in Fig. S2 is not surprising and again, doesn't really help the theme you wish to establish by focusing on the role of ROCK in skeletogenic cells over time. If anything, the data reported in Fig. S2 shows that perturbation of VEGF signaling has very little effect embryo-wide, while Fig. 1 shows that perturbation of VEGF signaling has a noticeable effect on ROCK expression in skeletogenic cells. If you choose to keep Fig. S2, I recommend that you indicate that embryo-wide vs skeletogenic cell difference more succinctly than given at present. It will also strengthen your paragraph in lines 110-127.

The importance of the western blot presented in Fig. S2 is to validate that the antibody recognizes a protein of the expected size. This strengthen the credibility of this commercial antibody to detect the sea urchin ROCK protein. We agree with the reviewer that the fact that the skeletogenic cells are less than 5% of the embryonic cells is important to explain why we didn’t see an affect of VEGFR inhibition in the western blot, and we changed the text to express it (lines 108-111): “Yet, this measurement was done on proteins extracted from whole embryos, of which the skeletogenic cells, where VEGFR is active, are less than 5% of the total cell mass (42). We therefore wanted to study the spatial expression of ROCK and specifically, its regulation in the skeletogenic cells.”

(3) Comparison of Fig. 2 and Fig. S3. To me the reader is confused when Fig. S3 is 33hpf as reported in the text (but not in the figure legend), and Fig. 2 shows 2 day old embryos - on the figure and figure legend but not in the text. So, the reader sees the text indicating 33hpf and looks around and the figure 2 says 2dpf. Does that mean 33hpf = 2dpf, the reader is thinking. To clarify, I suggest including the 2dpf in the text or simply drop the time in the text and report it in the two figures. Further, in the middle of the paragraph 130-143 you switch from reporting on Fig.S3 to Fig. 2, yet the reader doesn't know that. The reader is still looking at Fig. S3. The problem here is that at 33hpf the skeleton doesn't yet show the reduction or abnormalities that are shown later at 2dpf in Fig. 2. In clarifying this paragraph both the reduction in ROCK expression and the subsequent alterations in growth and patterning of the skeleton will be clear to the reader.

Thank you for raising this point. We added in the caption of Fig. S3 that the measurements were done in 33hpf. We also added in the text, that the observations of the skeletogenic phenotypes were done at 2dpf (48hpf). We made a break between the first paragraph discussing Fig. S3 and the paragraph discussing Fig. 2.

(4) The experiment with Y27632, an inhibitor of ROCK, is significantly improved in this revision. The concern earlier was the possibility that at the concentration used there might be off-target effects since other kinases are affected by higher concentrations of this selective inhibitor. The authors have modified this component of the paper and performed experiments at lower concentrations where other reports indicate the inhibitor is highly selective for ROCK, and they still demonstrate an inhibition of skeletal production. This, plus the added citations greatly increases confidence that this inhibition is selective for ROCK, thus enabling a stronger conclusion that ROCK has a role in skeletal growth and patterning.

Thank you for asking us to test this lower concentration which improved the credibility of our findings.

Line 239 - should be: indicating instead of indicting We corrected that.

(5) Line 402-403."The first step in generating the sea urchin spicules is the construction of the spicule cavity, a membrane filled with calcium carbonate and coated with F-actin (Fig. 8A)". I suggest more precise language. The way this now reads (above) is that somehow the spicule cavity is a membrane and that membrane is filled with CaCO3. And further the membrane is coated with F-actin. Isn't the spicule cavity what is filled with CaCO3? And isn't that cavity surrounded by a membrane? And the F-actin must be in the cortex of the cell since there is very little cytoplasm associated with the pseudopodial extensions that surround the spicule.

We change this sentence to: “The first step in generating the sea urchin spicules is the construction of the spicule cavity where the mineral is engulfed in a membrane coated with F-actin” (lines 403-404). Our observations show that F-actin is enriched around the spicule cavity. It could be an extension of the cell cortex, but we did not prove it, so we prefer to simply describe what we saw.

Line 405-408. Thank you for putting in this unknown. It is important to point out that while you've shown that ROCK contributes to regulation of actomyosin, it is not clear whether this is direct or indirect. You have also shown that ROCK somehow contributes to regulation of the GRN that leads to skeletogenesis. Thus, your data are consistent in showing that ROCK perturbation cripples normal skeletogenesis both via morpholino and with a selective inhibitor. Your last part of the discussion then offers speculation as to what might be affected specifically. That discussion sets the stage for digging even deeper to identify specific targets of ROCK activity.

Thank you, we agree with you that there is an exciting road ahead of us!

Version 3

Abstract

Abstract

Biomineralization had apparently evolved independently in different phyla, using distinct minerals, organic scaffolds and gene regulatory networks (GRNs). However, diverse eukaryotes from unicellular organisms, through echinoderms to vertebrates, use the actomyosin network during biomineralization. Specifically, the actomyosin remodeling protein, Rho-associated coiled-coil kinase (ROCK) regulates cell differentiation and gene expression in vertebrates’ biomineralizing cells, yet, little is known on ROCK’s role in invertebrates’ biomineralization. Here we reveal that ROCK controls the formation, growth and morphology of the calcite spicules in the sea urchin larva. ROCK expression is elevated in the sea urchin skeletogenic cells downstream of the Vascular Endothelial Growth Factor (VEGF) signaling. ROCK inhibition leads to skeletal loss and disrupts skeletogenic gene expression. ROCK inhibition after spicule formation reduces spicule elongation rate and induces ectopic spicule branching. Similar skeletogenic phenotypes are observed when ROCK is inhibited in a skeletogenic cell culture, indicating that these phenotypes are due to ROCK activity specifically in the skeletogenic cells. Reduced skeletal growth and enhanced branching are also observed under direct perturbations of the actomyosin network. We propose that ROCK and the actomyosin machinery were employed independently, downstream of distinct GRNs, to regulate biomineral growth and morphology in Eukaryotes.

Introduction

Biomineralization is the process in which organisms from the five kingdoms of life, use minerals to produce shells, skeletons and teeth that protect and support them (1–4). Recent studies suggest that biomineralization evolved by the phylum specific co-option of ancestral gene regulatory networks (GRNs) that drove the construction of distinct organic scaffolds (4–9) and by the evolution of specialized sets of biomineralization proteins (3, 7, 10–13). This explanation is in line with the dissimilar GRNs and biomineralization proteins that drive biomineralization in different phyla (5, 12, 14). Yet, there are common cellular processes required for mineralization in highly diverse phyla (4), suggesting that distinct upstream GRNs might have recruited common cellular and molecular mechanisms to drive biomineralization.

The actomyosin network was shown to play a role in biomineralization in various Eukaryote models, from unicellular organisms to vertebrates’ bones and teeth (15–22). There is a tight association between actin filaments and the biomineralization compartment in diatoms (19) and a highly dynamic actin organization that forms the biomineralization compartment in foraminifera (17). Perturbations of actin polymerization result with severely deformed shells and inhibition of shell shedding in coccolithophores (15, 16) and diatoms (18). In vertebrates, Rho GTPases and Rho-associated coiled-coil kinase (ROCK), regulate chondrocytes, osteoblasts and odontoblasts differentiation and affect genes expression in these biomineralizing cells (20–22). However, the roles of actomyosin remodeling in controlling mineral deposition and shaping biomineral morphology, are still unclear.

Sea urchin larval skeletogenesis provides an attractive model to study the role of the actomyosin network in biomineralization. Sea urchin larval skeletons are made of two frameworks of interconnected calcite rods, termed “spicules” that are generated by the skeletogenic cells (6, 23). To make the spicules, the skeletogenic cells form a ring with two lateral skeletogenic clusters and fuse through their filopodia forming a pseudopodia cable that links them into a syncytium (24, 25). The mineral is concentrated in the form of amorphous calcium carbonate (ACC) inside intracellular vesicles (26, 27). The vesicles are secreted into the biomineralization compartment generated in the skeletogenic cell clusters forming two triradiate spicules (Fig. 8A, (26, 28)). The tubular biomineralization compartment, also called the spicule cavity, elongates within the pseudopodia cable by localized mineral deposition at the tip of each rod (29).

The GRN that controls sea urchin skeletogenesis is known in great detail and is highly similar to the GRN that controls vertebrates’ vascularization, suggesting a common evolutionary origin of these two tubulogenesis programs (6, 23, 30). As the spicules elongate, the expression of key regulatory and biomineralization related genes becomes restricted to the skeletogenic cells proximal to the growing tips, possibly to promote mineral deposition at these sites (31–33). Localized gene expression is regulated by signaling cues such as the Vascular Endothelial Growth Factor (VEGF) signaling (30–32, 34). However, how the skeletogenic GRN drives spicule formation and localized mineral deposition at the tips of the rods is poorly understood.

Previous works suggest that the actomyosin network is essential for sea urchin skeletal growth. Calcium bearing vesicles perform an active diffusion motion in the skeletogenic cells with a diffusion length that inversely correlates with the strength and activity of the actomyosin network (35). Actin filaments are formed around the spicule (35, 36) and F-actin signal is enriched at the tips of the growing skeletal rods in skeletogenic cell culture (37). Genetic and pharmacological perturbations of the GTPase, CDC42, prevent the formation of filopodia in the skeletogenic cells and inhibit spicule formation and elongation (38). Pharmacological perturbations of ROCK prevent spicule formation (39) and genetic perturbations of Rhogap24l/2 result with ectopic spicule splitting (6). Despite these characterizations, little is known on the role of the actomyosin machinery in regulating sea urchin biomineral growth and morphology.

Here we study the role of ROCK and the actomyosin network in the sea urchin Paracentrotus lividus (P. lividus). Our findings reveal the critical role of ROCK and the actomyosin network in multiple aspects of sea urchin biomineralization, suggesting a common use of these factors in Eukaryote biomineralization, downstream of distinct GRNs.

Results

ROCK is enriched in the skeletogenic cells depending on VEGF signaling

We sought to study the spatial expression of the ROCK protein and its regulation by VEGF signaling, a prominent regulator of sea urchin skeletogenesis (6, 30). The sequence of ROCK and especially, its functional domains are highly conserved between human and sea urchin (Fig. S1, (40–42)). According to RNA-seq data measured in P. lividus, ROCK is a maternal gene that degrades after the maternal to zygotic transition and picks up again after hatching (>14 hours post fertilization (hpf), Fig. S2A). We used a commercial antibody that recognizes human ROCK protein to test ROCK expression in P. lividus embryos. We first used western blot to detect ROCK expression in control and under VEGFR inhibition at the time of spicule formation and during skeletal elongation (Axitinib treatment, ∼22hpf, 27hpf and 33hpf, Fig. S2D). The antibody detected a ∼150kDa protein (Fig. S2D) that is the predicted size for P. lividus ROCK protein (153kDa). VEGFR inhibition marginally increased the overall level of ROCK at 22hpf, but did not affect it at 27hpf and 33hpf (Fig. S2E). Yet, this measurement was done on proteins extracted from whole embryos, of which the skeletogenic cells, where VEGFR is active, are less than 5% of the total cell mass (43). We therefore wanted to study the spatial expression of ROCK and specifically, its regulation in the skeletogenic cells.

We studied the spatial distribution of ROCK protein using ROCK antibody and the skeletal cell marker, 6a9, that marks the skeletogenic syncytium membrane, in control and under VEGFR inhibition ((44), Fig. 1). We quantified ROCK signal in the skeletogenic cells compared to the neighboring ectodermal cells in both conditions, at the three time points mentioned above. In the three time points, under VEGFR inhibition, the skeletogenic cells do not migrate to their proper positions in the anterlateral and post-oral chains, as previously reported (6, 30, 34). At 22hpf, ROCK expression in the skeletogenic cells is mildly enriched compared to the neighboring ectodermal cells and this enrichment is not observed under VEGFR inhibition (Fig. 1A-F). At 27hpf, ROCK enrichment in the skeletogenic cells increases, but similar enrichment is observed under VEGFR inhibition (Fig. 1G-L). At 33hpf, ROCK enrichment in the skeletogenic cells is most apparent (Fig. 1M-O) and depends on VEGF signaling (Fig. 1P-S). At both 27hpf and 33hpf in control embryos, ROCK seems localized near the skeletogenic cell membranes, which further supports its activation in these cells, since ROCK activation leads to its localization to the cell membranes in other systems (45). Overall, this data demonstrate that ROCK expression is elevated in the skeletogenic cells and this enrichment strengthens with skeletal elongation and depends on VEGF signaling.

ROCK enrichment in the skeletogenic cells increases with time and depends on VEGF signaling.

(A-R) ROCK immunostaining at different times points in control and under VEGFR inhibition (150 nM Axitinib). In each condition and time point, the left image shows ROCK immunostaining and the skeletogenic cell marker, 6a9 (A, D, G, J, M and P). The middle image shows ROCK immunostaining alone in the whole embryo (B, E, H, K, N and Q). The right image shows the enlargement of the white rectangle region of the middle image (C, F, I, L, O, R). Scale bar in whole embryo is 50µm and in enlargements is 10µm. E – ectoderm, SC – skeletogenic cells. (S) quantification of the ratio between ROCK signal/area in the skeletogenic cells compared to the ectodermal cells (see methods for details). Each box plot shows the average marked in x, the median, the first and the third quartiles (edges of boxes) and all experimental measurement (dots). Experiments were performed in three independent biological replicates and in each condition at least 33 embryos were measured. Statistical significance was measured using paired 2-tailed t-test where, * indicates P<0.05, ** indicates P<0.005, and *** indicates P<0.0005.

ROCK activity in the skeletogenic cells controls spicule initiation, growth and morphology

Next, we studied the role of ROCK in sea urchin skeletogenesis using genetic perturbations. We downregulated the expression of ROCK by the injection of two different translation morpholino anti-sense oligonucleotides (MASOs). ROCK MASO-1 matches the start of translation region, ROCK MASO-2 matches the 5’ region upstream the start of translation, and random MASO was used as a control. Embryos injected with either ROCK MASO-1 or MASO-2 show reduced ROCK signal at 33hpf, supporting the downregulation of the ROCK protein using these MASO’s (Fig. S3B, F, J, N, R, V).

Embryos injected with ROCK MASO-1 or MASO-2 show strongly reduced skeletons at two days post fertilization (2dpf, Fig. 2C, D) and embryos injected with ROCK MASO-2 show an additional phenotype of severely branched skeleton at this time (Fig. 2B). The additional branching phenotype and the larger percentage of affected embryos in ROCK MASO-2 are probably due to its higher efficiency resulting from its more 5’ location relative to translation start codon (Fig. 2E). Thus, the genetic perturbations of ROCK expression indicate that sea urchin ROCK is important for skeletal elongation and normal branching pattern.

ROCK activity is essential for spicule formation, normal elongation and branching in whole embryos and in skeletogenic cultures.

(A-E) genetic perturbation of ROCK translation using two different MASOs observed at 2dpf. (A) control embryo injected with Random MASO. (B) embryo injected with ROCK MO-2 shows ectopic spicule branching. (C, D) embryos injected with ROCK MO-1 or MO-2 show spicule initiations. (E) summary of MASO injection phenotypes based on 4-6 independent biological replicates. (F-K) pharmacological perturbations of ROCK activity using 10µM and 80µM of the inhibitor Y27632 observed at 2dpf. (F) representative control embryo with normal skeletal rods, B, body; AL, anterolateral; PO, post-oral and MV, midventral. (G) complete skeletal loss in embryo treated continuously with 10µM ROCK inhibitor. (J) reduced skeletal growth and enhanced ectopic branching in embryo where 10µM ROCK inhibitor was added at 25hpf. (J) small spicules with enhanced ectopic branching in embryo where 80µM ROCK inhibitor was added at 25hpf. (K) summary of perturbation phenotypes based on 3-8 biological replicates for each treatment. See additional concentrations, phenotypes and summary in Fig. S4 and Table S1. (L-O) representative spicules from skeletogenic cell cultures in control and under 30µM Y27632 at 72hpf. (L) linear spicule in control culture, (M) Y27632 addition at 16hpf, before spicule initiation, completely blocks spiculogenesis. (N, O) Y27632 addition after spicule initiation, at 48hpf, reduces spicule elongation (N), and enhances branching (O). (P) quantification of spicule length in control and ROCK inhibition (>48hpf) at 72hpf. *P < 0.05, **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment, except from 120µm that was done in two biological repeats. Scale bars are 50µm. In L, N and O, the numbers at the bottom indicate the number of spicules that show this phenotypes (left) over all observed spicules (right).

To elucidate the dependence of ROCK phenotypes on ROCK activity level and identify its effect on different stages of skeletogenesis, we tested the skeletogenic phenotypes of different concentrations of ROCK inhibitor, Y27632, applied at different time points (Fig. S4A). Y27632 binds to the ATP site of ROCK’s kinase domain and prevents its activity (46). Y27632 affinity to ROCK is 100 times higher than its affinity to other kinases, such as PKA and PKC (47). The amino-acids to which Y27632 binds, are conserved in the sea urchin ROCK protein, supporting its specific inhibition of this protein (Fig. S1C). Y27632 had been used in cell cultures and live embryos from vertebrates to Drosophila, in concentrations between 10-100µM (47–49). In the sea urchin embryo a concertation of 75µM of Y27632 was reported to completely block skeleton formation (39) and a concentration of 100µM was shown to delay gut invagination (50). Therefore, we tested the range of 10-80µM Y27632 applied before or after spicule initiation (Fig. S4A).

Continuous ROCK inhibition beginning at egg fertilization, resulted with significant skeletogenic phenotypes that are dose dependent (Fig. 2, Fig. S4). Continuous ROCK inhibition using 80µM Y27632, did not affect skeletogenic cell migration but eliminated skeletal formation at 27hpf (Fig. S4B, C). At 2dpf, complete skeletal loss is detected under continuous ROCK inhibition in all concentrations, ranging from 15% of the embryos exposed to 10µM to 67% of the embryos exposed to 80µM Y27632 (Fig. 2G, K, in agreement with (39)). The rest of the embryos exposed to continuous ROCK inhibition show either spicule initiations or small spicules with enhanced ectopic branching, with percentage depending on Y27632 concertation (Fig. 2K, S4K). Except from the skeletogenic phenotypes, the overall embryonic development of embryos exposed to Y27632 in all concentrations seems normal (Fig. 2G, J). Similar results are observed when adding the inhibitor at 20hpf, with distribution of phenotypes that depends on the concentration (Fig. S4I, K). Importantly, skeletal loss, spicule initiation and ectopic branching were not observed under PKC or PKA inhibition, that resulted with much milder skeletogenic phenotypes (PKC) or no skeletogenic phenotype (PKA, (51, 52)), supporting the selective inhibition of ROCK by Y27632. Altogether, continuous ROCK inhibition results with severe skeletogenic phenotype ranging from complete skeletal loss to small spicule with ectopic branching, with ratios that depend on the inhibitor concentration.

The addition of the inhibitor at 25hpf, after spicule initiation, results in a majority of embryos showing ectopic spicule branching (Fig. 2H-K, S4F-H, K). The reduction of skeletal growth rate and ectopic branching can be observed a few hours after the addition of ROCK inhibitor (Fig. S5, 30µM). Washing the inhibitor at the highest concertation after 25hpf results in partial recovery of skeletogenesis, with normal or mildly delayed skeletons, indicating that ROCK inhibition is reversable (Fig. S4E, K).

Overall, the genetic and pharmacological perturbations of ROCK result with sever reduction of skeletal growth and enhanced skeletal branching, but only continuous ROCK inhibition leads to complete skeletal loss (Fig. 2G, K). Immunostaining of fertilized eggs clearly shows that ROCK protein is maternal, in agreement with our RNA-seq data (Fig. S2A-C). The injected MASO’s cannot interfere with the maternal ROCK protein whereas the inhibitor affects the activity of both maternal and zygotic ROCK, which could underlie the absence of skeletal loss in the genetic perturbations. Indeed, ROCK signal is only moderately reduced in ROCK MASO injected embryos at 33hpf (Fig. S3F, J, N), demonstrating the partial penetration of the MASO. Thus, our genetic and pharmacological perturbations reconcile, and indicate that ROCK activity is necessary for spicule formation, skeletal elongation and normal branching pattern.

To test if ROCK skeletogenic phenotypes are due to its activity specifically in the skeletogenic cells, we inhibited ROCK in a culture of isolated skeletogenic cells (Fig. 2L-P). The addition of ROCK inhibitor to the skeletogenic cell culture at 16hpf, before the spicules form, completely abolished spicule formation (Fig. 2M). The addition of the inhibitor at 48hpf, after spicule formation, resulted with significantly shorter spicules (Fig. 2L, N) and increased branching compared to control spicules (Fig. 2O). The reduction of the spicule length becomes more significant with increasing concentration of the ROCK inhibitor (Fig. 2P). Notably, skeletal loss and ectopic spicule branching were not observed in PKA or PKC inhibition in skeletogenic cell cultures (52), further supporting the selective inhibition of ROCK by Y27632. ROCK skeletogenic phenotypes in the isolated skeletogenic cell culture are similar to its phenotypes in whole embryos, verifying that ROCK activity in the skeletogenic cells is essential for spicule formation, normal elongation and prevention of branching.

ROCK inhibition reduces spicule volume, surface area and total length, but not thickness (SR-µCT)

To quantify the effect of ROCK inhibition on mineral deposition and skeletal shape we used synchrotron radiation micro-computed tomography (SR-µCT) that enables three-dimensional visualization and quantification of spicule morphology at 2 and 3dpf (Figs. 3, S6, 40µM Y27632, see methods for details). The aberrant morphology of ROCK inhibited spicules is visible at both developmental time points (Fig. 3A). The ectopic branching occurs both at the tips where tip-splitting is detected (arrowheads, Fig. 3A), as well as in the back, where small spicule extensions, or mineral ‘dripping’ are observed (arrow at 48hpf, Fig. 3A). This demonstrates that the regulation of mineral deposition is perturbed both at the tip and in the back of the spicules under ROCK inhibition.

SR-µCT studies of ROCK inhibited spicuels show reduction in skeletal volume, surface area and total length, but not thickness and 2-pulse calcein shows loss of tip-dominance.

(A) Exemplary 3D-renderings of control spicules (top, blue) and spicules where 40µM of ROCK inhibitor was added at 25hpf (bottom, orange), dissected at 48hpf and 72hpf. Arrowheads point to tip splitting and arrow at 48hpf point to spicule dripping at the back. (B) Spicule volume vs. area for control and ROCK-inhibited spicules at 48 and 72 hpf. Each data point represents a single spicule. (C-D) Frequency distributions for volume and surface area of control and ROCK-inhibited spicules at 48 hpf and 72 hpf (left and right violin plots, respectively). (E) Spicules’ total branch length and average thickness for control and ROCK-inhibited spicules at 48 and 72 hpf. (F-G) Frequency distributions for spicule lengths and thickness of control and ROCK-inhibited spicules dissected at 48hpf and 72hpf. (H, I) Calcein two-pulses experiment. Embryos were exposed to blue calcein at 29-30hpf, to green calcein at 44-45hpf and stained with FM4-64 membrane marker (red) prior to image acquisition at 48hpf. (H) Control embryo, (G) Embryo where 30µM of Y27632 was added at 25hpf. The experiments were done in three biological replicates and the numbers at the bottom indicate the number of embryos that show this phenotype out of all embryos scored.

When compared to control spicules, ROCK inhibition leads to a ∼2.5-fold reduction of the spicule volume and surface area in both 2 and 3dpf (Fig. 3B-D, see Table S2 for measured values and standard deviations and S3 for statistical analyses). The average rate of spicule growth between 48hpf and 72hpf in control embryos is 325.5µm3/hour, and reduces to 119.8µm3/hour under ROCK inhibition. Hence, the rate of mineral deposition is significantly reduced under ROCK inhibition.

To understand the specific effect of ROCK inhibition on spicule elongation vs. thickening, we compared the total length and the mean thickness of the spicules between control and ROCK inhibition (Fig. 3E-G). The total length of control spicules was on average about 2.5-times longer than the length under ROCK inhibition in both 2dpf and 3dpf (Fig. 3E, F, tables S2 and S3). In contrast to its effect on the spicule length, ROCK inhibition caused a minor increase of the spicule thickness at 2dpf and at 3dpf the thickness was not affected by the inhibition (Fig. 3E-G, Tables S2, S3). These SR-µCT measurements indicate that ROCK activity plays a crucial role in controlling the rate of mineral deposition at the tips of the rods and spicule elongation, and is not affecting spicule thickening.

ROCK activity does not control mineral uptake but is required for “tip-dominance”

To test the effect of ROCK inhibition on mineral intake, we used the fluorescent chromophore, calcein that binds to divalent ions including Ca+2 and is widely used to track the calcium pathway in biomineralization (6, 28, 33, 35). We measured the number of calcein stained vesicles per area as well as the number of pixels of calcein signal per area in the skeletogenic cells, in control and under ROCK inhibition at 1dpf and 2dpf (Fig. S7A-E). Continuous ROCK inhibition does not change the number of calcium bearing vesicles at 1dpf, but the number of calcein stained pixels significantly increases in this condition, possibly indicating that the vesicle volume is larger (Fig. S7F, G). At 2dpf, the number of calcein stained pixels is similar in control and in continuous ROCK inhibition, suggesting that the overall calcium content does not change in this condition (Fig. S7H, I). The number of calcein stained vesicles is however decreased under ROCK inhibition, possibly indicating that there are fewer vesicles with larger volume. Addition of ROCK inhibitor at 25hpf affects neither the number of calcein stained vesicles nor the number of calcein stained pixels in the skeletogenic cells. Together, these measurements show that ROCK is not required for the uptake of calcium into the cells. Therefore, the skeletal phenotypes occurring under ROCK inhibition are not related to a decrease in calcium uptake or shortage in intracellular calcium, but are due to defects in mineral processing within the cells and in mineral deposition.

To monitor the effect of ROCK inhibition on mineral deposition rate and distribution, we applied two pulses of different calcein dyes (53): blue calcein was added at 29hpf and washed after one hour, followed by green calcein added at 44hpf and washed after one hour. Hence, blue calcein marks the initial triradiate region and green calcein marks the edges of the spicules that were recently deposited (blue and green arrows in Fig. 3H, I). The dye FM4-64 was used to mark the cell membranes. Under ROCK inhibition, the green labeled region is much closer to the blue labeled region compared to control, due to the reduction in skeletal elongation rate, in agreement with our SR-µCT measurements. However, while in control embryos each body rod has a single tip stained in green calcein, under ROCK inhibition one of the body rods has two tips and both are stained in green. This indicates that mineral is being deposited in both edges, and the mechanism that prevents the growth of multiple tips in each rod, enabling “tip-dominance”, (in analogy to plant stem apical dominance), is disrupted under ROCK inhibition.

The activity of the actomyosin network is essential for normal spicule elongation and inhibition of ectopic branching

In other systems, major roles of ROCK are to control F-actin polymerization and MyoII activation (54–58), hence we wanted to directly test the role of these actomyosin components in sea urchin skeletogenesis. To directly inhibit F-actin polymerization we used Latrunculin-A (Lat-A), that prevents the polymerization of actin monomers (59) and was shown to be effective in sea urchin embryos (60). To directly inhibit actomyosin contractility we used Blebbistatin (Blebb), that interferes with MyoII ATPase activity, prevents actin-myosin interactions (61) and was shown to be effective in sea urchin embryos (62). To prevent interference with the early cell divisions where the actomyosin network plays a critical role (63–65), we added the inhibitors, individually or together, before or after spicule formation (20hpf and 25hpf, see methods for details). We tested the resulting phenotypes at 2dpf (Fig. 4A-G).

Actin polymerization and myosin activation affect skeletal growth and branching.

(A-F) representative embryos showing the effect of actomyosin perturbations at 2df. (A) control embryo (B) embryo treated with 2μM Blebbistatin >25hpf , (C) embryo treated with 2μM Blebbistatin >20hpf, (D) embryo treated with 2nM Latrunculin-A and 1.5μM Blebbistatin >25hpf, (E) embryo treated with 2nM Latrunculin-A >20hpf, (F) embryo treated with 2nM Latrunculin-A >25hpf, arrow pointing to the additional spicule rod. (G) Statistics of Latrunculin-A and Blebbistatin phenotypes, color code of phenotype is indicated in the representative images. Error bars indicate standard deviation. All treatments were conducted in at least three biological replicates and exact number of replicates and scored embryos are provided in table S4. (H-J) representative spicules recorded at 72hpf from (H) control skeletogenic cell culture, (I) skeletogenic cell culture were 2nM Latrunculin-A was added at 48hpf and (J) skeletogenic cell culture were 2μM Blebbistatin was added at 48hpf. (K) Quantification of spicule length in the different treatments at 72hpf **P < 0.001, Kruskal-Wallis non-parametric test. Results are based on three biological repeats for each treatment. Scale bars are 50μm.

Our results indicate that F-actin polymerization and to a lesser extent, actomyosin contractility, are essential for normal skeletal growth and inhibition of ectopic branching. The majority of the embryos where MyoII activation was inhibited show ectopic branching mostly at the tips of the rods, and a minor delay in skeletal growth (Fig. 4B, C, G). These phenotypes are quite similar to those observe in the addition of 10µM of ROCK inhibitor at 25hpf (Fig. 2H, K). The majority of the embryos where F-actin formation was inhibited show significant reduction of skeletal growth and severe ectopic skeletal branching (Fig. 4E, F and similar phenotypes to Fig. 4D, statistics in Fig. 4G). These phenotypes are quite similar to those observed in 10µM continuous inhibition of ROCK and the addition of 30µM of ROCK inhibitor at 20hpf (Fig. 2H, K). In some embryos, additional spicules form (Fig. 4F, G). The co-inhibition of F-actin polymerization and MyoII activity results in skeletogenic phenotypes that resemble, but are slightly stronger, than those of the inhibition of F-actin alone (Fig. 4D, G). Thus, the phenotypes resulting from the inhibition of MyoII activation resemble those of the late inhibition of ROCK at low inhibitor concertation. The phenotypes resulting from inhibition of actin polymerization resemble those of the continuous inhibition of ROCK in low inhibitor concertation and the late inhibition of ROCK at higher inhibitor concentrations, yet skeletal loss is only observed under ROCK continuous inhibition.

To test whether actin polymerization and myosin activity specifically in the skeletogenic cells underlies the observed skeletogenic phenotypes, we inhibit both in an isolated skeletogenic cell culture (Fig. 4H-K). We added the inhibitors after the spicule formed (>48hpf) and observed the phenotypes at 3dpf. Both inhibitors reduced skeletal growth, with Lantruculin-A having a significantly stronger effect than Blebbistatin (Fig. 4K), in agreement with the stronger skeletogenic phenotypes of Lantruculin-A in whole embryos. However, differently than in whole embryos, skeletal branching in both Lantruculin-A and Blebbistatin is not affected and is similar to the observed branching in control cultures. The absence of the branching phenotype in the skeletogenic cell culture could be due to the increased rigidity of the substrate that could compensate for the reduce actomyosin activity. Another option is that the branching phenotype in whole embryos is due to the reduction of the activity of the actomyosin network in non-skeletogenic cells. Nevertheless, these results indicate that normal skeletal growth depends primarily on F-actin polymerization and to a lesser extent, on myosin contractility in the skeletogenic cells.

ROCK activity is required for F-actin organization around the forming spicule

The similarity between the skeletogenic phenotypes under ROCK inhibition and the direct perturbations of the actomyosin network led us to test the effect of ROCK inhibition on F-actin organization and MyoII activity. To accomplish this we used phalloidin, a fluorophore that binds F-actin specifically, an antibody that detects phosphorylated MyoII (MyoIIP, (35)) and the skeletogenic marker, 6a9 (Fig. 5). In control embryos, F-actin is detected around the tri-radiate calcite spicules at 27hpf (green arrow in Fig. 5B, in agreement with (35, 36)). The pseudopodia cable that connects the skeletogenic cells shows lower F-actin signal compared to the signal around the spicule (blue line in Fig. 5B, C, E blue arrows in Fig. 5G, H, J). MyoIIP signal is not enriched in the skeletogenic cells nor in the spicule (Fig. 5D, I). Under continuous ROCK inhibition, the pseudopodia that connects the skeletogenic cells still forms (Fig. 5R) and enhanced F-actin signal is still observed in the lateral skeletogenic cell clusters (green arrow in Fig. 5L); but unlike control embryos, the spicule cavity is not formed and F-actin is not organized around it (Fig. 5L). MyoIIP signal seems unchanged by ROCK inhibition at this time (Fig. 5N, S), in agreement with the weak and late phenotype of MyoII inhibition (Fig. 4). Thus, the pseudopodia cable forms but F-actin organization around the spicules does not occur under continuous ROCK inhibition, however it is not clear if the effect on F-actin organization is direct or due to the absence of spicule in this condition.

ROCK inhibition effect on F-actin organization and MyoII activity at 27hpf.

Representative images showing normal embryo (A-J) and embryos treated with ROCK inhibitor from fertilization (K-T, 80µM), at 27hpf. Phalloidin (green) was used to stain F-actin, MyollP (red) was used to stain phosphorylated myosinII and 6a9 (blue) was used to mark the skeletogenic cells. Right panels show enlargements of the rectangle sections marked in the left most panels. Green arrows point to regions that show enriched F-actin signal, blue arrows point to the pseudopodia cable. The blue line in B, C and E marks a region of the pseudopodia cable that is stained by 6a9 but has low F-actin signal. The experiments were done in 3 biological replicates, the numbers at the bottom left of (A, K) indicate the number of embryos that show this phenotype out of all embryos scored. Scale bar in A, F, K and P is 50μm, and in B, I, N and S is 10μm.

F-actin is enriched at the tips of the elongating spicules independently of ROCK activity

To assess the role of ROCK in actomyosin organization during skeletal elongation we compared F-actin and myoIIP signals between control embryos and embryos where ROCK inhibitor was added at 25hpf (Fig. 6). At the prism stage, we did not detect a clear difference in MyoIIP signal in the skeletogenic cells, between control and ROCK inhibited embryos (Fig. 6D, I). At this time, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Figs. 6B, G) as well as in ROCK morphants (green arrowheads in Fig. S3K, W). In both control and ROCK inhibited embryos, the F-actin signal was markedly higher in the regions of the pseudopodia cable where the spicule cavity had formed, compared to the regions where the spicule cavity had not yet formed (blue arrows in Fig. 6B-E, G-J). We quantified the F-actin signal at the tips compared to the back in control and ROCK inhibited embryos (Fig. 6K, L, see methods for details). The phalloidin signal is on average 3-fold stronger at the tips of the spicules compared to the back, in both control and ROCK inhibition, indicating that F-actin enrichment at the tips is significant and independent of ROCK activity.

F-actin filaments are enriched at the tips of the spicules and the tip to back ratio is unaffected by ROCK inhibition

(A-J) representative images at 33hpf, showing normal embryo (A-E) and embryos treated with 30µM ROCK inhibitor from the gastrula stage (>25hpf, F-J). Embryos are stained with Phalloidin (green), MyollP antibody (red) and 6a9 (blue). (B-E, G-J) enlargement sections of the spicule area marked with rectangle in A and F. White arrowheads point to the enriched F-actin signal at the tips. Blue arrows point to the region of the pseudopodia cable that is not filled with the spicule cavity. (K, L) quantification of the tip to back F-actin signal (number of green pixels per area) at 33hpf in control embryos and ROCK inhibition >25hpf. Each box plot shows the average (x), median (middle line), the first and the third quartiles and all the points measured. Asterisks indicate statistical significance in paired t-test, where * is p<0.05 and *** is p< 0.0005, NS-not significant. (M-V) similar experiments to (A-J), at 48hpf, 40µM of ROCK inhibitor added at 25hpf. Red arrows point to non-skeletogenic cells enriched with MyoIIP. The experiments were repeated in 3 biological replicates and the numbers at the bottom left of (A, F, M, R) indicate the number of embryos that show this phenotype out of all embryos scored. Scale bar in A, F, M, R is 50µm and in B, G, N, S, is 10µm. (W, X) representative spicules out of three biological replicates from control skeletogenic cell culture (W, n=60), and skeletogenic cell treated with 30µM ROCK inhibitor added at 48hpf and recorded at 72hpf (X, n=52). Left panel is phase image, middle panel is phalloidin staining and right panel shows the overlay. Green arrows point to the enhanced F-actin signal at the tips. Scale bar is 20μm.

At the pluteus stage, F-actin is enriched at the tips of the spicules in both control and ROCK inhibited embryos (white arrowheads in Fig. 6N-Q, S-V). Some non-skeletogenic cells that are enriched with MyoIIP signal are detected at this time (red arrowheads in Fig. 6N-Q). Together these data demonstrate that F-actin coats the spicule cavity and is enriched at the tip of the rods, independently of ROCK activity.

We tested the effect of ROCK inhibition on F-actin in skeletogenic cell cultures. F-actin is detected at the tips, in both control cultures and cultures where ROCK was inhibited after spicule formation (Fig. 6W, X). Yet, under ROCK inhibition, branching is enhanced and F-actin is enriched at the splitting tips of the spicule rods (Fig, 6X). The enrichment of F-actin at the splitting tips of the spicule under ROCK inhibition in skeletogenic cell culture resembles the calcein staining at the two tips in the embryo in this condition (Fig. 3I). These observations further support the role of ROCK activity in regulating tip-dominance during sea urchin skeletogenesis.

ROCK activity is essential for normal skeletogenic gene expression

The role of ROCK in regulating gene expression during vertebrates’ chondrocytes, osteoblasts and odontoblasts differentiation (20–22) intrigued us to study the effect of ROCK inhibition on skeletogenic gene expression in the sea urchin embryo. ROCK continuous inhibition significantly downregulates the expression of multiple skeletogenic genes, including the cytoskeleton remodeling gene, rhogap24l/2, the biomineralization genes, caral7, scl26a5 and SM30 at 27hpf and 2dpf (Fig. 7A, (6)). ROCK inhibition did not affect the expression level of VEGFR and ROCK itself at these times (Fig. 7A). Washing the inhibitor after 25hpf or adding the inhibitor at 25hpf had a weaker effect, but still led to downregulation of caral7, SM30, angio1 and notch1 (Fig. 7B). Thus, ROCK activity is required for the normal expression level of multiple skeletogenic genes during sea urchin skeletogenesis.

ROCK activity is essential for normal gene expression in the skeletogenic cells.

(A, B) the effect of 80µM ROCK inhibition on gene expression. (A) continuous ROCK inhibition at 27hpf and 50 hpf (n=4). (B) addition of ROCK inhibitor at 25hpf and the wash of ROCK inhibitor at 25hpf, measured at 50hpf (n=3). Asterisks indicate p<0.001, one tailed z-test. Error bars show standard deviation. (C-F) Representative images of control embryo (top panels), embryos where ROCK inhibitor was added at 25hpf (middle panels), and embryos that were exposed to continuous ROCK inhibition (bottom panels), at the pluteus stage (∼48hpf). (C) skeletogenic cell marker, 6a9. MV, midventral; AL, anterolateral, and PO, Post-oral rods, D, dorsal skeletogenic chain. (D-F) spatial expression of skeletogenic genes. Gene names are indicated at the top of each panel. Numbers at the bottom of each image indicate the number of embryos that show this phenotype (left) out of all embryos scored (right), conducted in at least three independent biological replicates.

The role of ROCK and the actomyosin network in sea urchin skeletogenesis, summary.

(A) the spicule forms at about 24hpf in P. lividus and at 27hpf, the triradiate spicule is coated with F-actin. (B) ROCK activity is required for spicule initiation and for F-actin organization around the spicule. (C) model of the functional links between the skeletogenic GRN, ROCK, F-actin and their skeletogenic outcomes, during spicule initiation (∼24-27hpf in P. lividus). Cian arrows indicate links discovered in this work. (D) during skeletal elongation F-actin is detected around the spicule cavity and is enriched at the tips of the rods. The expression of SM50 and some other skeletogenic genes, is localized to the tips of the rods. (E) under the addition of ROCK inhibitor, skeletal growth rate is reduced and ectopic spicule branching is observed. The expression of skeletogenic genes is localized to the vicinity of the growing rods. (F) the functional links between the skeletogenic GRN, ROCK, F-actin and MyoII and their skeletogenic outcomes during skeletal elongation. Cian arrows indicate links discovered in this work.

Both continuous and late ROCK inhibition strongly affect the spatial expression of key regulatory and biomineralization genes that are normally expressed differentially within the skeletogenic lineage at the pluteus stage (Figs. 7D-F, S8). These include, the VEGF receptor (VEGFR), the biomineralization genes SM50 and SM30, skeletogenic transcription factors, Ets1, Alx1 and Erg1 that are essential for skeletogenic cell specification and biomineralization (23, 66) and the transcription factors Hex and MyoD1. In normal plutei these genes are differentially expressed within skeletogenic lineage, but in both continuous and late ROCK inhibition the expression of these genes becomes localized to the two lateral skeletogenic cell clusters and their vicinity (Fig. 7D-F, S8). The lateral skeletogenic cells clusters are the cells where these genes are expressed in the gastrula stage (24-27hpf) and where spicule formation begins in normal embryos (6, 32). The expression of VEGFR and SM50 under ROCK inhibition is more expanded than the other tested genes and is also observed at the anterolateral rods (Fig. 7D, E). The VEGF ligand that is normally expressed in the ectodermal cells near the four tips of the post-oral and anterolateral rods (32), is expressed at four ectodermal domains near the lateral skeletogenic cell clusters and near the end of the anterolateral chain, under ROCK inhibition (Fig. S8G). Our data shows that under ROCK inhibition, the expression of key skeletogenic genes remains localized to the area near the lateral skeletogenic cells clusters (Fig. 7D-F, S8), despite the proper formation of the dorsal, anterolateral and mid-ventral chains (Fig. 7C). Overall, ROCK activity is essential for normal expression level and spatial gene expression in the skeletogenic lineage.

Discussion

Biomineralization is a complex morphogenetic process, encoded in the species genome and executed by the GRNs and the cellular machinery. Here we used the sea urchin larval skeletogenesis to investigate the role of ROCK and the actomyosin network in biomineral formation and in regulating gene expression, as we summarize in Fig. 8. ROCK expression is enriched in the skeletogenic cells, partially depending on VEGF signaling (Fig. 1, 8C). ROCK activity in the skeletogenic cells is necessary for spicule formation, skeletal elongation and prevention of ectopic skeletal branching (Figs. 2-3, S4, S5). Direct inhibition of F-actin polymerization results in similar but not identical skeletogenic phenotypes (Fig. 4). F-actin is organized around the spicule cavity and enriched at the spicule tips (Figs. 5, 6). ROCK activity feeds back into the skeletogenic GRN and affects regulatory and biomineralization gene expression (Fig. 7). Below we discuss our findings and their implications on the biological control and evolution of biomineralization.

The first step in generating the sea urchin spicules is the construction of the spicule cavity where the mineral is engulfed in a membrane coated with F-actin (Fig. 8A). ROCK activity is necessary for the spicule initiation and for the organization of F-actin around the spicule cavity (Figs. 2, 5, 8B). However, since the spicule doesn’t form under ROCK inhibition, it is hard to conclude if the absence of F-actin organization around the spicule is due to ROCK regulation of actin polymerization or due to the absence of spicule in this condition. Relatedly, a direct perturbation of F-actin polymerization strongly disrupts spicule growth and shape, but not spicule formation, indicating that F-actin is not essential for spicule initiation (Fig. 4). We therefore propose that ROCK activity is necessary for spicule initiation, but not through its’ direct regulation of the actomyosin network, but though its other targets, yet to be identified (Fig. 8C).

After the spicule cavity has formed, the spicules elongate inside the pseudopodia chord by mineral deposition at the tips of the spicule rods, which depends on ROCK activity, on F-actin polymerization and to a lesser extent, on MyoII contractility (Figs. 2-4, 8D-F). F-actin is detected along the spicules and is significantly enriched at the tips of the elongating spicules, independently of ROCK activity (Fig. 5, 6, 8D, E (67)). Under ROCK inhibition and under the inhibition of F-actin polymerization, ectopic branching is observed at the tips and in the back (Figs. 2H, J, O, 3A, I, 4, 8E). MyoII inhibition results with ectopic branching at the tips of the rods (Fig. 4B, G), that is similar to the effect of late ROCK inhibition at low concertation (Fig. 2H, K). These correlative similarities between ROCK and the actomyosin perturbations lead us to the following speculations: the low dosage of late ROCK inhibition is perturbing mostly ROCK activation of MyoII contractility while the higher dosage affects factors that control actin polymerization (Fig. 8F). Further studies in higher temporal and spatial resolution of MyoIIP activity and F-actin structures in control and under ROCK inhibition will enable us to test this.

The participation of ROCK, F-actin and MyoII in polarized growth and vesicle exocytosis has been observed in both animals and plants (67, 68). F-actin is used for short range myosin trafficking of vesicles toward the secretion site (67, 69). Once the secretory vesicle fuses with the membrane, it is coated with F-actin that stabilizes the vesicle and enables MyoII binding and ROCK-regulated contraction, necessary for content secretion (48, 49). In plant cells, F-actin accumulate at the growing tip of the cell and assists vesicle exocytosis necessary for polarized tip-growth (68). ROCK, F-actin and MyoII could be regulating the localized exocytosis of mineral-bearing vesicles during sea urchin skeletogenesis. The reduction of spicule growth rate and the enhanced branching could be due to impaired and misplaced vesicle trafficking and deposition under ROCK and F-actin inhibition. Detailed investigation of the kinetics and deposition of mineral-bearing vesicles and the role of actomyosin network in these processes will illuminate the regulation of vesicle exocytosis during biomineralization.

ROCK activity is necessary for correct spatiotemporal expression of regulatory and biomineralization genes (Fig. 7). ROCK inhibition downregulates the expression level of skeletogenic genes (Fig. 7A) and prevents the spatial progression of key skeletogenic genes toward the edges of the skeletal rods (Fig. 7, S8, 8E). This implies that ROCK activity provides regulatory cues to the skeletogenic GRN required for the dynamic progression of gene expression within the skeletogenic lineage. In other systems, ROCK activity was shown to regulate gene expression through various intracellular pathways (70–72). Furthermore, ROCK has an important role in mechano-sensing of matrix rigidity and transducing it into changes in gene expression (73). Relatedly, stiff substrate activates ROCK in pre-osteoblastic cells in-vitro, which activates Erk that leads to osteogenic differentiation and upregulation of osteogenic gene expression (72). Sea urchin ROCK could be a part of the mechano-sensing mechanism that detects the high spicule stiffness and transduces it into gene expression, providing a necessary feedback between spicule elongation and the skeletogenic GRN.

Overall, our findings together with the role of the actomyosin network in biomineralization, from calcifying and silicifying single cells to vertebrates’ bones and teeth (15–22, 72), suggest that this molecular machinery is a part of the common molecular tool-kit of biomineralization. Most likely, the actomyosin network was employed independently, downstream of distinct GRNs across Eukaryotes, to control mineral growth and morphogenesis.

Methods

Animal and embryos

Adult Paracentrotous lividus were obtained from the Institute of Oceanographic and Limnological Research (IOLR) in Eilat, Israel. Eggs and sperm were obtained by injection of 0.5M KCl solution into the adult sea urchin species. Embryos were cultured at 18°C in 0.2µ filtered ASW.

Imaging

Embryonic phenotypes and WMISH were imaged by Zeiss Axioimager 2. Fluorescent markers were imaged using Nikon A1-R Confocal microscope. All images were aligned in Adobe photoshop and Illustrator.

VEGFR inhibitor, Axitinib (AG013736) treatment

Axitinib (AG013736, Selleckchem, Houston, TX, USA) was applied as described in (32)

Western Blot

Embryo cultures treated with 150 nM Axitinib, or DMSO as control, were collected at 22hpf, 27hpf and 33hpf by centrifugation of 10,000g for 2 minutes. We lysed the pellet in 150µL lysis buffer (20 mM Tris-HCl, 150 mM NaCl, 1% Triton X-100, pH 8) with protease inhibitors (protease inhibitor cocktail w/o metal chelator; Sigma, P8340; 1:100) and phosphatase inhibitors (50 mM NaF, 1 mM Na3VO4, 1 mM Na4P2O7, 1mM, βGP) as described in (74). 40µg protein were loaded on 8% SDS-acrylamide gel, transferred to PVDF membranes for Western blot analysis. Anti-ROCK antibody (ab45171, Abcam) was used in a 1:300 dilution followed by anti-rabbit HRP secondary antibody diluted to 1:2500 (Jackson ImmunoResearch 111-035-003). For the loading control, the membrane was incubated with anti-β-tubulin antibody (1:5000; Sigma) followed by anti-mouse antibody 1:5000 (Jackson ImmunoResearch 115-035-003). Quantification of ROCK signal was done using Image studio lite vr. 5.2. ROCK signal was divided by Tubulin signal and then the ratio between ROCK/Tubulin in control vs. VEGFR inhibition was calculated. The graph in Fig. S2 was generated based on three biological replicates (different sets of parents) at 27hpf and 33hpdf and four biological replicates at 22hpf.

Immunostaining procedure

Phalloidin labeling, MyoIIP (p20-MyosinII), ROCK (anti ROCK2+ROCK1 antibody [EP786Y], AB-ab45171, Abcam), and 6a9 immunostaining were done similarly to (35).

Quantification of Anti-ROCK and Phalloidin signal

A graphical program was written to allow manual quantification of the fluorescent signal from the spicule regions by identifying “stained” pixels per selected area. The ratio of stained anti-ROCK to region total area was compared between ectodermal and skeletogenic cells, and between control and VEGFR inhibited embryos at 22, 27 and 33hpf. The average ratios and ztest for significance difference from 1, are 22hpf control, ∼1.3, z=0.0006; VEGFR inhibition ∼1, z=0.5, 27hpf control ∼1.8, z<10-11; VEGFR inhibition ∼1.6, z<10-8, 33hpf control ∼4.6, z<10-23; VEGFR inhibition ∼1.5, z<10-5. For phalloidin analyses, embryos at 33hpf of control and ROCK inhibitor addition at 25hpf were used and for each image, two regions were selected: an area at the tip of the spicule and an area at the back of the spicule, about ∼10 microns apart. The ratio of stained phalloidin to region total area was compared between tip and back, in control and ROCK inhibited embryos. Paired t-test was used to compare the differences in phalloidin signal between the tip and back of the spicules across all groups and between control and ROCK inhibited embryos.

ROCK MASOs injections

Microinjections were conducted similarly to (32). The eggs were injected with an injection solution containing 0.12M KCl, 0.5µg/µl Rhodamine Dextran and 900µM of MASOs. Translation MASOs were designed and synthesized by Gene Tools, Philomath, OR, according to ROCK sequence (40). The sequence of ROCK MASO-1 (5’-AGACATATTTGGAGCCGA[CAT]CCTG-3’) matches the start of translation region and ROCK MASO-2 (5’-TCTCTTGCGTTATATTCCACTAAGT-3’) matches the 5’ region upstream of the start of translation (40). Control embryos were injected in the same concentration with Random MASO. Injected embryos were cultured at 18°C, imaged and scored at 2dpf.

ROCK inhibitor (Y27632) treatment

Y27632 stock (10005583 CAS Regisrty No. 129830-38-2, Cayman chemical), was prepared by reconstituting the chemical in DMSO. The embryos were treated with Y27632 at a final concentration between 30-80µM, as mentioned in the results. Throughout the paper, control embryos were cultured in DMSO at the same volume as Y27632 solution and no more than 0.1% (v/v).

ROCK inhibition in isolated skeletogenic cell culture

Skeletogenic cell culture was performed as described in (37) with minor changes. Isolated skeletogenic cells were cultured in CultureSlides (4 chambers polystyrene vessel tissue culture treated glass slide REF 354104) with ASW (Red Sea Fish LTD) containing Gentamicin and Penicillin-Streptomycin (GPS) at 18°C. Media containing 4% (v/v) horse serum (Sigma-Aldrich H1270) in ASW+GPS was added to the culture. ROCK inhibitor, Y27632, was added to the cell culture at 16hpf (before spicule initiation) or 48hpf (after spicule initiation), and images were taken at 72hpf. Each treatment was conducted in three biological repeats except from the >48hpf addition of 120µM which was done in two biological repeats.

Quantification of skeletal length and statistical analysis

Skeletal length measurement was done as described in (32). The measurements were repeated for three biological repeats for control, ROCK inhibition with 30μM and 80μM, Latrunculin-A 2nM, Blebbistatin 2µM, and 2 biological repeats for 120μM. In ROCK inhibition experiments, a total of 275 skeletons were measured for control, 105 skeletons for 30μM, 122 skeletons for 80μM and 80 for 120μM. In Latrunculin-A and Blebbistatin experiments, a total of 116 skeletons were measured for control, 149 for Latrunculin-A and 107 for Blebbistatin. The data were analyzed in Excel and the statistical analysis was performed using Kruskal-Wallis non-parametric test in Excel as described in (75) and SPSS statistics 27.

Spicule preparation for micro-computed tomography (SR-µCT) measurement

Control embryos and embryos where 40µM Y27632 was added at 25hpf, were collected by centrifugation of 50mL Falcon tubes at 2000rpm, at 48h and 72hpf. Every tube was washed in 10 mL cold distilled water (DW), transferred into epi-tube and washed three times in 1mL cold DW (incubation at 4°C between every wash, centrifuge at 2000rpm in room temperature for 2min). Then the skeletons were washed three times in 3% NaOCl followed by three washes in DW, one wash in 70% EtOH and one wash in Acetone. Skeletal spicules were dried at room temperature and stored at -20 °C.

Synchrotron and lab-based SR-µCT

Dry and loose skeletal spicules prepared as described above, were used for the SR-µCT analysis. We analyzed 14 sets of spicules from three (control) and four (ROCK inhibited) pairs of parents, divided into four groups: 2dpf, control, n=44; 2dpf ROCK inhibited, n=93; 3dpf control, n=51; 3dpf ROCK inhibited, n=93, see Table S1). Spicules were glued to sharpened toothpick tips, either in bulk or individually when picked by hand by means of a brush bristle. Each toothpick loaded with spicules, was mounted on a 3ml vial twist-off lid and for transportation securely stored in the vial (Fig. S5A, B). For each set 2-3 µCT-samples were prepared, which allowed the acquisition of statistically relevant sample sizes (40+ spicules/group, Table S1). Tomographic data sets were acquired using the following scanning parameters; for the lab based µCT (RX Solutions, Chavanod, France) 90kV, 143-160µA, 1440 images, average frames 10, frame rates 0.4-1, voxel sizes: 0.45-59µm; and for the synchrotron radiation µCT (Anatomix, Synchrotron SOLEIL, Gif-sur-Yvette, France) 2048 images, angle between projections 0.09°, voxel size: 0.65µm. Tomographic data reconstruction was performed with commercial (RX Solutions, Chavanod, France) and PyHST2 software (A. Mirone et al., Nucl. Instrum. Meth. B 324 (2014) 41-48, doi:10.1016/j.nimb.2013.09.030) at the MPICI and at Anatomix, SOLEIL, respectively.

Three-dimensional (3D) data analysis

Data visualization, pre-processing including cropping and image type conversion to 16bit (of the Sr-µCT data), and user-augmented segmentation of intact spicules were performed in Amira using off-the-shelf modules (Fig. S5C-E, (76)). Volume and area measurements were performed on the segmented spicules, i.e. the label-fields using Amira’s material.statistics module (Volume3D and Area3D). Average spicule thickness and total length measurements were performed with python code, utilizing available python libraries (e.g., skimage, skan, scipy). The code was applied to 3D-tif files created from Amira’s label-fields, which contained the segmented spicules. For length measurements, the spicules were skeletonized and the chord lengths of the branches forming a spicule-skeleton were summed up to obtain the total spicule length. For thickness measurements, we performed Euclidean distance transforms for each spicule using scipy’s ndimage.morphology.distance_transform_edt package, and calculated the average spicule thickness based on the distance measurements obtained at each voxel from spicule-skeleton to the nearest spicule surface. Quality control of our measurements was performed with open source software, Fiji and available plugins (e.g. BoneJ, FiberJ, DiameterJ).

Calcein staining

A 2mg/ml stock solution of blue calcein (M1255, Sigma) and a 2mg/ml stock solution of green calcein (C0875, Sigma) were prepared by dissolving the chemicals in distilled water. Working solution of 250μg/ml and 150μg/ml, respectively, was prepared by diluting the stock solution in artificial sea water. Blue calcein was added to the embryo culture at 29hpf for one hour and then washed. At 44hpf the green calcein was added for one hour and washed. For FM4-64 staining A 100µg/ml stock solution of FM4-64 (T13320, Life technologies, OR, USA) was prepared by dissolving the chemical in distilled water. Working solution of 2.5 µg/ml was prepared by diluting the stock solution in artificial sea water. Embryos were immersed in working solution about 10 minutes before visualization.

Latrunculin-A and Blebbistatin treatments

1mM of Lat-A stock (L12370, Thermo Fisher Scientific) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Lat-A at final concentration of 2 nM. 1mM of Blebbistatin stock (13013-5, Enco) was prepared by reconstituting the chemical in DMSO. The embryos were treated with Blebbistatin at final concentration of 2 µM. For the double inhibition experiments, embryos were treated with Blebb at final concentration of 1.5 µM, and with Lat-A at final concentration of 2 nM.

WMISH procedure

WMISH was performed as described in (32). List of the primers used to make the WMISH probes is available in (32).

cDNA preparation for QPCR experiments

For ROCK inhibition experiments total RNA was extracted from >1000 sea urchin embryos in each condition using RNeasy Mini Kit (50) from QIAGEN (#74104) according to the kits’ protocol. DNase treatment on column was done using RNase-Free DNase Set-Qiagen (50) (#79254). RNA was reverse-transcribed using the High Capacity cDNA RT kit, AB-4368814 (Applied Biosystems) following the manufacturer’s instructions.

Quantitative polymerase chain reaction (QPCR)

QPCR was carried out as described in (32). Complete list of primer sequences used for the experiments in provided there.

Supporting information